### 2017-02-13 paper.tex
update

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                          template.tex
%
% LaTeX template for papers conforming to the United States Sections of
% the Combustion Institue style guide.
%
% Authors:
%     Bryan W. Weber, University of Connecticut
%     Kyle E. Niemeyer, Oregon State University
%
% This work is licensed under the Creative Commons Attribution 4.0
% International License. To view a copy of this license, visit
% http://creativecommons.org/licenses/by/4.0/.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{ussci}

%======================================================================
\usepackage{todonotes}
\usepackage{graphicx}
\usepackage[binary-units]{siunitx}
\usepackage{gensymb}
\sisetup{group-separator={,},
	detect-all,
	binary-units,
	list-units = single,
	range-units = single,
	range-phrase = --,
	per-mode = symbol-or-fraction,
	separate-uncertainty = true,
	multi-part-units = single,
	list-final-separator = {, and }
	%    scientific-notation = fixed
}
%======================================================================
% Add your bibliography file here, replace template.bib
\addbibresource{paper.bib}
%======================================================================
% Replace "Reaction Kinetics" in the line below by your paper topic
\newcommand\papertopic{Reaction Kinetics}
%======================================================================

\title{ Title of Paper }

\author[1]{Author Name}
\author[1]{Author Name}
\author[2]{Author Name}
\author[2,*]{Author Name}

\affil[1]{Department, Institution, Address, Country}
\affil[2]{Department, Institution, Address, Country}
\affil[*]{Corresponding author: \email{author@university.edu}}

\begin{document}
\maketitle

%====================================================================
\begin{abstract} % not to exceed 200 words
Abstract should be between 150--200 words and should state briefly the purpose
of the research, the principal results and major conclusions. An abstract is
often presented separately from the article, so it must be able to stand alone.
For this reason, References should be avoided, but if essential, then cite the
author(s) and year(s). Also, non-standard or uncommon abbreviations should be
avoided, but if essential they must be defined at their first mention in the
abstract itself.
\end{abstract}

% (Provide 2-4 keywords describing your research. Only abbreviations firmly
% established in the field may be used. These keywords will be used for
% sessioning/indexing purposes.)
\begin{keyword}
    Keyword1\sep Keyword2\sep Keyword3\sep Keyword4
\end{keyword}

%====================================================================
\section{Introduction}
%

Single Instruction, Multiple-Data (SIMD) processing, also known as vector-processing, is an important parallel computing paradigm that is increasingly being used in scientific computing.
While central processing unit (CPU) clock speeds have increased rapidly over the past few decades---the phenomena known as Moore's Law---power consumption and head dissipation issues have caused this trend to slow.
Traditional multi-core parallelism is often used to increase CPU performance, however SIMD processors---and the related Single Instruction, Multiple Thread (SIMT) processors, e.g. graphics processing units (GPUs)---have become more popular due to their increased floating operation throughput.
The parallel programming standard OpenCL~\cite{stone2010openc} has further enabled adoption of vector-processing based codes in scientific computing by providing a common application program interface (API) for execution on heterogeneous systems (e.g. CPU, GPU, etc.).

\begin{figure}[h]
	\centering
	\includegraphics[width=0.33\linewidth]{SIMD.pdf}
	\caption{Schematic of SIMD processing.  A single processing element (e.g. a CPU core) contains a vector unit with several lanes (L0, L1, etc.).  The vector unit executes a single instruction concurrently on multiple data. Adapted from\cite{Simdfig:2016}}
	\label{F:SIMD}
\end{figure}

A SIMD instruction utilizes specialized vector processing hardware to execute the same floating point operation (e.g. multiplication, division, etc.) on multiple pieces of data concurrently (Fig.~\ref{F:SIMD}).
Nearly all modern CPUs have vector processing units capable of running SIMD instructions with typical vector widths of \SIrange{128}{256}{\bit}, e.g. the SSE or AVX2 instruction sets.
Additionally, specialized hardware accelerators, e.g. Intel's Xeon Phi co-processor (or MIC), have been developed that have tens of cores with very wide (\SIrange{256}{512}{\bit}) SIMD units---the AVX-512 instruction set---while forthcoming Intel CPUs (the Skylake Xeon and Cannon Lake architectures) are also expected to support AVX-512.
\todo[inline]{Find a source for this?  It all seems to be weird cpu journalism sites, maybe search intel}

Several recent works, e.g. \cite{CurtisGPU:2017,Sewerin20151375} have investigated SIMT-based chemical kinetic integration on GPUs, however SIMD-based chemical kinetics have been less studied in comparison.
\textcite{stone2016} implemented a linearly-implicit fourth-order stiff Rosenbrock solver in the OpenCL for various platforms including the CPU, GPU and the MIC; the Rosenbrock solver performance was compared to that of a non-stiff, fourth-order explicit Runge–Kutta–Fehlberg integrator.
Two parallelization models were investigated in OpenCL, the so called \textit{shallow}-SIMD vectorization, where each SIMD-lane solved a single chemical kinetic initial value problem (IVP), and a SIMT-vectorization where each thread solved a single IVP.
SIMT-vectorized chemical source term evaluation was \SI{13}{\percent} and \SI{35}{\percent} slower than the equivalent OpenMP code on the CPU and MIC respectively.
The SIMD-vectorized chemical source terms were up to \SI{3}{$\times$} faster on both platforms than the OpenMP baseline; GPU chemical source term evaluation was of similar performance to the SIMD-vectorized versions.
SIMD-vectorization improved the integrator performance over the OpenMP baseline by \SIrange{2.5}{2.8}{$\times$} on the CPU and \SIrange{4.7}{4.9}{$\times$} on the MIC.
\textcite{kroshko2013efficient} implemented a shallow-SIMD -vectorized third order Rosenbrock solver (RODAS) integrator for atmospheric chemistry on a Cell Broadband Engine---a specially designed vector processor---finding a speedup of \SI{1.89}{$\times$} over a serial version of the same code, a \SI{94}{\percent} parallel efficiency.

This work will study the performance of chemical source term evaluation of automatically generated shallow-SIMD-vectorized codes for a wide range chemical kinetic models on the CPU.
The performance will be compared to a baseline SIMT-vectorized codes to determine the effective SIMD speedup.
Finally future extensions to this work will be detailed.
\todo[inline]{and OpenMp, time permitting}

\section{Methodology}
\subsection{Chemical Kinetics Equations}
The current code evaluates the following chemical source terms:

\begin{equation}
\frac{\text{d}[C_k]}{\text{d}t} = \sum_{i}^{N_R} \nu_{k,i} R_i c_i
\end{equation}
where $\frac{\text{d}[C_k]}{\text{d}t}$, $N_R$, $\nu_{k,i}$, ${R_i}$, $c_i$ are the time rate of change of the concentration of species $k$, the number of reactions in the model, the net stoichiometric coefficient of species $k$ in reaction $i$, the net rate of progress of reaction $i$, and the pressure modification of reaction $i$ respectively.
For further detail on each term, the reader is referred to our previous work~\cite{Niemeyer:2016aa}; the code is capable of evaluating all modern reaction rate types (e.g. pressure-log, chebyshev, etc.).

In addition, the temperature rate of change using the constant-pressure assumption\footnote{Note: this refers to evaluation within a reaction sub-step in the operator splitting scheme.} is evaluated as:
\begin{equation}
\frac{\text{d}[T]}{\text{d}t} = -\frac{\sum_{k}^{N_S} H^{\degree}_k \dot{\omega}_k}{\sum_{k}^{N_S} [C_k] c_{p_k}^{\degree}}
\end{equation}
although the code is equally capable of a constant-volume assumption.

\subsection{Code Generation}
Code generation is handled by the python package \texttt{loopy}~\cite{kloeckner_loopy_2014}, which provides an easily accessible

\section{Conclusions}
%

\section{Acknowledgements}
This research was funded by \ldots

\noindent\textbf{Page Limits:} The total length of the paper including references should be limited to 6 pages.

\printbibliography[heading=bibintoc]

\end{document}

### 2017-02-13 paper.tex
updates

@@
 graphics processing units (GPUs)---have become more popular due to their increased floating operation throughput.
-The parallel programming standard OpenCL~\cite{stone2010openc} has further enabled adoption of vector-processing based codes in scientific computing by providing a common application program interface (API) for execution on heterogeneous systems (e.g.
+The parallel programming standard OpenCL~\cite{stone2010opencl} has further enabled adoption of vector-processing based codes in scientific computing by providing a common application program interface (API) for execution on heterogeneous systems (e.g.
 CPU, GPU, etc.).
 
-\begin{figure}[h]
+\begin{figure}[ht]
 	\centering
@@
 \end{equation}
-although the code is equally capable of a constant-volume assumption.
+although the code is equally capable of a evaluating the temperature rate using a constant-volume assumption, we omit this here for brevity.
 
 \subsection{Code Generation}
-Code generation is handled by the python package \texttt{loopy}~\cite{kloeckner_loopy_2014}, which provides an easily accessible
+Code generation is handled by the python package \texttt{loo.py}~\cite{kloeckner_loopy_2014}, which provides an interface to OpenCL, while allowing for unit testing and simple changes of program structure, e.g.
+data ordering, vectorization, threading patterns etc.
+Additionally, \texttt{loo.py} allows for code generation for multiple languages, including (non-vectorized) C, CUDA, and Intel's ISPC compiler, which targets SIMD vectorization on the CPU and MIC.
+This is a direction that may be explored in the future to compare the speedups of the various compilers.
+
+\subsection{Data Ordering and Vectorization Patterns}
+For a chemical kinetic model with species $k$ for the $j$-th thermo-chemical state to be evaluated, $[C_k]_j$ there are two choices for data-storage in memory.
+In the ``C'' (row-major) format, species concentrations for $j$-th thermo-chemical state are stored sequentially in memory, i.e.: $[C_0]_0\text{,}\allowbreak C[1]_0\ldots\allowbreak [C_{N_S}]_0\text{,}\allowbreak [C_0]_1\allowbreak\ldots\allowbreak [C_{N_S}]_N$, where $N$ is the total number of thermo-chemical conditions.
+Alternatively, in the ``F'' (Fortan column-major) format species concentrations a single species for all thermo-chemical states are sequential, i.e.: $[C_0]_0\text{,}\allowbreak[C_0]\ldots\allowbreak[C_0]_N\text{,}\allowbreak[C_1]_0\ldots\allowbreak[C_{N_S}]_N$
+In Sec.~\ref{results} the effect of this ordering will be shown to have a large performance impact on the program runtime.
+
+In this work, a \textit{shallow}-SIMD vectorization has been implemented, entailing each SIMD-lane evaluating the species rates for an thermo-chemical state independently.
+In future efforts, a so called \textit{deep}-SIMD vectorization will be pursued allowing vector-units to cooperate to evaluate species rates for a single thermo-chemical state.
+A deep-vectorization may result in SIMD-\textit{waste} in this context---similar to thread-divergence in SIMT-processing---caused by different SIMD-lanes executing different instructions (e.g.
+from differing for-loop bounds), however past studies demonstrate the merits of deep-SIMT vectorization on GPUs~\cite{Sewerin20151375}.
+
+\section{Results and Discussion}
+
+\label{results}
 

### 2017-02-14 paper.tex
update validation

@@
 \usepackage{gensymb}
+\usepackage{amsmath}
 \sisetup{group-separator={,},
@@
 
-In addition, the temperature rate of change using the constant-pressure assumption\footnote{Note: this refers to evaluation within a reaction sub-step in the operator splitting scheme.} is evaluated as:
+In addition, the temperature rate of change using the constant-pressure assumption\footnote{Note: in this context, the ``constant-pressure assumption'' refers to evaluation within a reaction sub-step in the operator splitting scheme, rather than a general constant-pressure reactive-flow simulation.} is evaluated as:
 \begin{equation}
@@
 Alternatively, in the ``F'' (Fortan column-major) format species concentrations a single species for all thermo-chemical states are sequential, i.e.: $[C_0]_0\text{,}\allowbreak[C_0]\ldots\allowbreak[C_0]_N\text{,}\allowbreak[C_1]_0\ldots\allowbreak[C_{N_S}]_N$
-In Sec.~\ref{results} the effect of this ordering will be shown to have a large performance impact on the program runtime.
+In Sec.~\ref{S:results} the effect of this ordering will be shown to have a large performance impact on the program runtime.
 
@@
 \section{Results and Discussion}
+\subsection{Validation}
+The reaction rates of progress, species and temperature rates in this study are validated by comparison to Cantera~\cite{Cantera}, however due to the intricacies of floating point arithmetic care must be taken when comparing results.
+
+In obtaining the net rate of progress of reaction $i$ from the forward and reverse rates of progress, $R_{i} = R_{i}^{\prime} - R_{i}^{\prime\prime}$, a base-10 bound on the loss of precision can be found as~\cite{goldberg1991every}:
+\begin{equation}
+2^{-p_i} \le \left\lvert 1 - \frac{R_{i}^{\prime}}{R_{i}^{\prime\prime}} \right\rvert \le 2^{-q_i}
+\label{e:bounds}
+\end{equation}
+with $q_i \le p_i$, while $1 - \frac{R_{i}^{\prime}}{R_{i}^{\prime\prime}}$ can be computed with accurately via use of higher-precision numbers (in this case, via use of Python's Decimal class).
+
+Plugging the forward and reverse rate constants from Cantera into Eq.~\eqref{e:bounds} we can obtain a reasonable a-priori precision bounds $\text{P}_{\text{min},i} = 2^{-p_i}$, $\text{P}_{\text{max},i} = 2^{-q_i}$.
+Finally, when determining the error in the net rates of progress between this code and Cantera, the range of percent error attributable to precision loss, $\text{P}_{\text{err},i}$ can be found as:
+\begin{equation}
+\frac{\left\lvert\ R_{i,\text{CT}} - R_{i}\right\rvert}{\text{P}_{\text{min},i}} \le
+\frac{\text{P}_{\text{err},i}}{100} \le \frac{\left\lvert R_{i,\text{CT}} - R_{i} \right\rvert}{\text{P}_{\text{max},i}}
+\label{e:rel_bound}
+\end{equation}
+where the \text{CT} subscript indicates values from Cantera.
+Note that the differences in Eq.~\eqref{e:rel_bound} are not subject to the same scrutiny as when computing the net rate of progress, namely because if they were small enough to merit this analysis the error would be minimal anyways.
+
+With this we can properly validate the code against Cantera, as seen in Table~\todo[inline]{this}
+
+\subsection{Results}
+\label{S:results}
 

### 2017-02-14 paper.tex
refine intro

@@
 
-Single Instruction, Multiple-Data (SIMD) processing, also known as vector-processing, is an important parallel computing paradigm that is increasingly being used in scientific computing.
+Single Instruction, Multiple-Data (SIMD) processing, also known as vector-processing, is an important parallel computing paradigm used increasingly in scientific computing.
 While central processing unit (CPU) clock speeds have increased rapidly over the past few decades---the phenomena known as Moore's Law---power consumption and head dissipation issues have caused this trend to slow.
 Traditional multi-core parallelism is often used to increase CPU performance, however SIMD processors---and the related Single Instruction, Multiple Thread (SIMT) processors, e.g.
-graphics processing units (GPUs)---have become more popular due to their increased floating operation throughput.
+graphics processing units (GPUs)---have gained recognition due to their increased floating operation throughput.
 The parallel programming standard OpenCL~\cite{stone2010opencl} has further enabled adoption of vector-processing based codes in scientific computing by providing a common application program interface (API) for execution on heterogeneous systems (e.g.
@@
 multiplication, division, etc.) on multiple pieces of data concurrently (Fig.~\ref{F:SIMD}).
-Nearly all modern CPUs have vector processing units capable of running SIMD instructions with typical vector widths of \SIrange{128}{256}{\bit}, e.g. the SSE or AVX2 instruction sets.
+Nearly all modern CPUs have vector processing units capable of running SIMD instructions, typically capable of handling 2--4 concurrent double precision operations---known as the vector width.
 Additionally, specialized hardware accelerators, e.g.
-Intel's Xeon Phi co-processor (or MIC), have been developed that have tens of cores with very wide (\SIrange{256}{512}{\bit}) SIMD units---the AVX-512 instruction set---while forthcoming Intel CPUs (the Skylake Xeon and Cannon Lake architectures) are also expected to support AVX-512.
-\todo[inline]{Find a source for this?  It all seems to be weird cpu journalism sites, maybe search intel}
+Intel's Xeon Phi co-processor (or MIC), have been developed that have tens of cores with very wide vector widths (e.g.
+4--8 double precision operations); these very wide vector widths are also expected on forthcoming Intel CPUs (the Skylake Xeon and Cannon Lake architectures).
 

### 2017-02-14 paper.tex
more edits

@@
 Two parallelization models were investigated in OpenCL, the so called \textit{shallow}-SIMD vectorization, where each SIMD-lane solved a single chemical kinetic initial value problem (IVP), and a SIMT-vectorization where each thread solved a single IVP.
-SIMT-vectorized chemical source term evaluation was \SI{13}{\percent} and \SI{35}{\percent} slower than the equivalent OpenMP code on the CPU and MIC respectively.
-The SIMD-vectorized chemical source terms were up to \SI{3}{$\times$} faster on both platforms than the OpenMP baseline; GPU chemical source term evaluation was of similar performance to the SIMD-vectorized versions.
-SIMD-vectorization improved the integrator performance over the OpenMP baseline by \SIrange{2.5}{2.8}{$\times$} on the CPU and \SIrange{4.7}{4.9}{$\times$} on the MIC.
+Shallow-SIMD vectorization improved the integrator performance over the OpenMP baseline by \SIrange{2.5}{2.8}{$\times$} on the CPU and \SIrange{4.7}{4.9}{$\times$} on the MIC, while the GPU performance while the GPU was \SIrange{1.4}{1.6}{$\times$} slower than the OpenMP baseline due to thread-divergence concerns.
 \textcite{kroshko2013efficient} implemented a shallow-SIMD -vectorized third order Rosenbrock solver (RODAS) integrator for atmospheric chemistry on a Cell Broadband Engine---a specially designed vector processor---finding a speedup of \SI{1.89}{$\times$} over a serial version of the same code, a \SI{94}{\percent} parallel efficiency.
@@
 Finally future extensions to this work will be detailed.
-\todo[inline]{and OpenMp, time permitting}
 
@@
 \subsection{Code Generation}
-Code generation is handled by the python package \texttt{loo.py}~\cite{kloeckner_loopy_2014}, which provides an interface to OpenCL, while allowing for unit testing and simple changes of program structure, e.g.
+Code generation is handled by the python package \texttt{loo.py}~\cite{kloeckner_loopy_2014}, which provides an interface to OpenCL allowing for unit testing and changes of program structure, e.g.
 data ordering, vectorization, threading patterns etc.
@@
 In the ``C'' (row-major) format, species concentrations for $j$-th thermo-chemical state are stored sequentially in memory, i.e.: $[C_0]_0\text{,}\allowbreak C[1]_0\ldots\allowbreak [C_{N_S}]_0\text{,}\allowbreak [C_0]_1\allowbreak\ldots\allowbreak [C_{N_S}]_N$, where $N$ is the total number of thermo-chemical conditions.
-Alternatively, in the ``F'' (Fortan column-major) format species concentrations a single species for all thermo-chemical states are sequential, i.e.: $[C_0]_0\text{,}\allowbreak[C_0]\ldots\allowbreak[C_0]_N\text{,}\allowbreak[C_1]_0\ldots\allowbreak[C_{N_S}]_N$
+Alternatively, in the ``F'' (Fortan column-major) format species concentrations a single species for all thermo-chemical states are sequential, i.e.: $[C_0]_0\text{,}\allowbreak[C_0]_1\ldots\allowbreak[C_0]_N\text{,}\allowbreak[C_1]_0\ldots\allowbreak[C_{N_S}]_N$
 In Sec.~\ref{S:results} the effect of this ordering will be shown to have a large performance impact on the program runtime.

### 2017-02-15 paper.tex
updates to manuscript

@@
 \usepackage{amsmath}
+\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}, e.g., \ce{H2SO4}
+\usepackage{booktabs,multicol} %better tables
 \sisetup{group-separator={,},
@@
 \begin{equation}
-\frac{\left\lvert\ R_{i,\text{CT}} - R_{i}\right\rvert}{\text{P}_{\text{min},i}} \le
-\frac{\text{P}_{\text{err},i}}{100} \le \frac{\left\lvert R_{i,\text{CT}} - R_{i} \right\rvert}{\text{P}_{\text{max},i}}
+\text{P}_{i,\text{min}} = 100 * \frac{\text{P}_{\text{min},i}}{\left\lvert\ R_{i,\text{CT}} - R_{i}\right\rvert} \le
+\text{P}_{i} \le 100 * \frac{\text{P}_{\text{max},i}}{\left\lvert R_{i,\text{CT}} - R_{i} \right\rvert} = \text{P}_{i,\text{max}}
 \label{e:rel_bound}
@@
 where the \text{CT} subscript indicates values from Cantera.
-Note that the differences in Eq.~\eqref{e:rel_bound} are not subject to the same scrutiny as when computing the net rate of progress, namely because if they were small enough to merit this analysis the error would be minimal anyways.
+The error of a quantity $X_{i,j}$ for state $j$, and index $i$ was determined as:
+\begin{equation}
+ E_{X_{i}} = \left\lVert \left\lVert \frac{\left\lvert X_{i,j,\text{CT}} - X_{i,j}\right\rvert}{\num{e-10} + \num{e-6} * \left\lvert X_{i,j,\text{CT}} \right\rvert} \right\rVert_{j,\infty} \right\rVert_{i, \infty}
+\label{e:rel_err}
+\end{equation}
+Finally, to estimate the effect of loss of precision on the net rates of progress, we let $P_{\text{min}}^{*} = P_{i,\text{min}}$ for the $i,j$ value that results in $E_{R_{i}}$.
 
-With this we can properly validate the code against Cantera, as seen in Table~\todo[inline]{this}
+\begin{table}[ht]
+\sisetup{retain-zero-exponent=true}
+\centering
+\begin{tabular}{@{}S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1] S[table-format=1.2e1] @{}}
+\toprule
+\multicolumn{1}{c}{Model} & \multicolumn{1}{c}{$E_{R_{i}^{\prime}}$} & \multicolumn{1}{c}{$E_{R_{i}^{\prime\prime}}$} &\multicolumn{1}{c}{$E_{R_{i}}$} & \multicolumn{1}{c}{$\text{P}_{\text{min}}^{*}$} & \multicolumn{1}{c}{$E_{\frac{\text{d}[C_k]}{\text{d}t}}$} & \multicolumn{1}{c}{$E_{\frac{\text{d}[T]}{\text{d}t}}$} \\
+\midrule
+\multicolumn{1}{c}{\ce{H2}\slash~\ce{CO}} & 1.20e-8 & 6.33e-8 & 1.32e1 & 4.66e6 & 2.37e1 & 2.12e5 \\
+\multicolumn{1}{c}{GRI-Mech.~3.0}  & 3.07e-8 & 5.90e-8 & 1.21e0 & 2.67e04 & 2.84e0 & 2.64e4 \\
+\multicolumn{1}{c}{USC-Mech II}  & 1.07e-7 & 1.37e-7 & 2.39e0 & 4.12e02 & 7.92e00 & 3.22e03 \\
+\bottomrule
+\end{tabular}
+\caption{Summary of rate of progress, species and temperature rate correctness.
+Error statistics are based on the infinity-norm of the relative error detailed in Eq.~\eqref{e:rel_err} for each quantity.
+}
+\label{T:error}
+\end{table}
+
+In Table~\ref{T:error}, we see the results of this code as compared to Cantera on a library of partially stirred reaction conditions described in our previous works~\cite{CurtisGPU:2017,Niemeyer:2016aa}
 

### 2017-02-15 scripts/data_parser.py
add first version of plotters and data

@@ -1,0 +1,39 @@
+import collections
+import os
+import json
+
+rundata = collections.namedtuple('data', 'num_conditions comptime overhead runtime')
+mechdata = collections.namedtuple('mechdata', 'mech num_species num_reactions')
+run = collections.namedtuple('run', 'lang vecwidth order vectype platform rates kernel cores rundata mechdata')
+
+def parse_data(directory='performance'):
+    #find all mechs
+    mechs = [x for x in os.listdir(directory)
+                if os.path.isdir(os.path.join(directory, x))]
+    perf_data = {}
+    for mech in mechs:
+        perf_data[mech] = []
+        path = os.path.join(directory, mech)
+        #get files
+        files = [os.path.join(path, x) for x in os.listdir(path)]
+        files = [x for x in files if os.path.isfile(x)]
+        # read mechanism info
+        mech_meta_file = next(x for x in files if x.endswith('.json'))
+        with open(mech_meta_file, 'r') as f:
+            mech_info = mechdata(**json.load(f))
+        files.remove(mech_meta_file)
+        for file in files:
+            #get name
+            name = os.path.basename(file)
+            #get run parameters
+            params = name[:name.index('_output')].split('_')
+            #get data
+            with open(file, 'r') as f:
+                lines = [[float(x) for x in l.strip().split(',')] for l in f.readlines() if l.strip()]
+            datalist = [rundata(*x) for x in lines]
+            #add to run
+            params.extend([datalist, mech_info])
+            perf_data[mech].append(
+                run(*params))
+
+    return perf_data

### 2017-02-15 scripts/error.py

@@ -1,0 +1,50 @@
+import os
+import numpy as np
+script_dir = os.path.dirname(os.path.abspath(__file__))
+error_path = os.path.join(script_dir, 'error_checking')
+mechs = [os.path.join(error_path, x) for x in os.listdir(error_path) if os.path.isdir(os.path.join(error_path, x))]
+
+err_dicts = {}
+for mech in mechs:
+    mech_name = os.path.basename(os.path.normpath(mech))
+    #get file list
+    files = [os.path.join(mech,x) for x in os.listdir(mech) if os.path.isfile(os.path.join(mech,x))]
+    err_dicts[mech_name] = {}
+    for file in files:
+        with open(file, 'r') as f:
+            lines = [x for x in f.read().split('\n') if x.strip()]
+        for line in lines:
+            name = line[:line.index(':')].strip()
+            err_arr = np.array([float(x) for x in line[line.index(':') + 1:].split(',') if x.strip()])
+            if ('rop_fwd' in name or 'rop_rev' in name) and 'linf' in name and np.any(err_arr > 1):
+                print file, np.where(err_arr > 1)
+            if name in err_dicts[mech_name]:
+                try:
+                    assert np.allclose(err_dicts[mech_name][name], err_arr)
+                    err_dicts[mech_name][name] = err_arr
+                except AssertionError:
+                    err_dicts[mech_name][name] = np.maximum(err_dicts[mech_name][name], err_arr)
+            else:
+                err_dicts[mech_name][name] = err_arr
+
+def format(val):
+    return '{:1.2e}'.format(val)
+for mech in err_dicts:
+    print(mech)
+    for name in err_dicts[mech]:
+        if 'l2' in name:
+            continue
+        err_vals = err_dicts[mech][name][np.where(np.logical_not(
+            np.isnan(err_dicts[mech][name])))]
+        if 'wdot' in name:
+            print('tdot', format(err_vals[0]))
+            print('species', format(np.linalg.norm(err_vals[1:], ord=np.inf)))
+        elif 'rop_net ' in name:
+            #find prevision range
+            maxv = np.linalg.norm(err_vals, ord=np.inf)
+            maxind = np.where(err_dicts[mech][name] == maxv)[0][0]
+            print(name, format(maxv))
+            print('precision_max', format(err_dicts[mech]['rop_net_precmax'][maxind]))
+            print('precision_min', format(err_dicts[mech]['rop_net_precmin'][maxind]))
[…]

### 2017-02-15 scripts/general_plotting.py

@@ -1,0 +1,146 @@
+"""Module with
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import plot_styles as ps
+
+font_size = 'large'
+
+legend_key = {'H2':r'H$_2$/CO',
+              'CH4':r'GRI-Mech 3.0',
+              'C2H4':r'USC-Mech II',
+              'IC5H11OH':r'IC$_5$H$_{11}$OH'
+              }
+
+def plot(plotdata, to_plot, minx=None, miny=None, plot_std=True,
+         norm={}, return_y=False, reacs_as_x=False, labels=[], plot_ind=None
+         ):
+    """Plot performance as a function of reaction count.
+    """
+    if reacs_as_x:
+        plotdata = sorted(plotdata, key=lambda x: x.mechdata.num_reactions)
+        x_vals = [x.mechdata.num_reactions for x in plotdata]
+    else:
+        plotdata = sorted(plotdata, key=lambda x: x.mechdata.num_species)
+        x_vals = [x.mechdata.num_species for x in plotdata]
+    for i, plot in enumerate(to_plot):
+        y_vals = []
+        err_vals = []
+        for run in plotdata:
+            if plot in norm:
+                run_y = [np.array(getattr(data, plot)) / data.num_conditions for data in run.rundata]
+            else:
+                run_y = [getattr(data, plot) for data in run.rundata]
+            err_vals.append(np.std(run_y))
+            y_vals.append(np.mean(run_y))
+
+        if plot_ind is not None:
+            assert labels
+            marker, hollow = ps.marker_wheel[plot_ind]
+            color = ps.color_wheel[plot_ind]
+            name = labels[plot_ind]
+        else:
+            marker, hollow = ps.marker_dict[plot]
+            color = ps.color_dict[plot]
+            name = plot
+
+        argdict = {'x':x_vals,
+                   'y':y_vals,
+                   'linestyle':'',
+                   'marker':marker,
+                   'label':ps.pretty_names(name)
+                   }
+        argdict['color'] = color
+        argdict['markeredgecolor'] = color
+        if not hollow:
+            argdict['markersize'] = ps.marker_style['size']
+        else:
+            argdict['markerfacecolor'] = 'None'
+            argdict['markersize'] = ps.clear_marker_style['size']
+
+        if plot_std:
+            argdict['yerr'] = err_vals
+        if plot_std:
+            line = plt.errorbar(**argdict)
+        else:
[…]

### 2017-02-15 scripts/perf_plotter.py

@@ -1,0 +1,120 @@
+import data_parser
+import argparse
+import plot_styles as ps
+import matplotlib.pyplot as plt
+import general_plotting as gp
+
+run = data_parser.run
+rundata = data_parser.rundata
+data = data_parser.parse_data()
+reacs_as_x = False
+norm = None
+
+def plotter(plot_name='plot.pdf', **filters):
+    #apply filters
+    for f in filters:
+        if filters[f] is None:
+            continue
+        assert any(data[x] for x in data), 'No data matching all filters'
+        if f in data[data.keys()[0]][0]._asdict():
+            for mech in data:
+                data[mech] = [x for x in data[mech] if getattr(x, f) == filters[f]]
+    #compute data
+    for mech in data:
+        for run in data[mech]:
+            rund = run.rundata
+    #now plot data
+    to_plot = ['runtime']
+    if filters['plot_compilation']:
+        to_plot.append('comptime')
+    if filters['plot_overhead']:
+        to_plot.append('overhead')
+    #get data
+    plot_data = [x for mech in data for x in data[mech]]
+    #get labels
+    diff_check = [x for x in run._fields if x not in ['mechdata', 'rundata']]
+    diffs = [set([getattr(x, check) for x in plot_data]) for check in diff_check]
+    #get # with more than 1 option
+    diff_locs = [i for i in range(len(diffs)) if len(diffs[i]) > 1]
+    assert len(diff_locs) <= 1, "Don't know how to create graphs for more than one diff"
+    if not diff_locs:
+        #regular plot
+        gp.plot(plot_data, to_plot, norm=['runtime'])
+    else:
+        loc = diff_locs[0]
+        diffs = sorted(diffs[loc])
+        labels = [diff_check[loc] + ' ' + x for x in diffs]
+        for i, val in enumerate(diffs):
+            match = [x for x in plot_data if getattr(x, diff_check[loc]) == val]
+            gp.plot(match,
+                to_plot, norm=['runtime'], labels=labels, plot_ind=i)
+    plt.ylabel(r'Runtime ($\frac{\si{\milli\second}}{\text{state}}$)')
+    plt.xlabel(r'Number of Species in Model')
+    plt.legend(**ps.legend_style)
+    ps.finalize()
+    plt.show()
+
+
+if __name__ == '__main__':
+    parser = argparse.ArgumentParser()
+    parser.add_argument('--lang',
+        required=False,
+        default='opencl',
+        type=str)
[…]

### 2017-02-15 scripts/plot_styles.py

@@ -1,0 +1,57 @@
+#a single consolidated place to import
+#such that all figures have identical styling (when possible)
+
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+import os.path
+
+#setup latex
+plt.rc('text', usetex=True)
+plt.rc('font', family='serif')
+plt.rc('text.latex',
+    preamble=r'\usepackage{amsmath},\usepackage{siunitx}')
+plt.rc('font', family='serif')
+
+legend_style = {'loc':0,
+    'fontsize':16,
+    'numpoints':1,
+    'shadow':True,
+    'fancybox':True
+}
+
+tick_font_size = 20
+
+marker_style = {
+    'size' : 15
+}
+
+clear_marker_style = {
+    'size' : 17
+}
+
+default_keys = ['runtime', 'comptime', 'overhead']
+marker_wheel = [('o', True), ('>', True), ('s', True)]
+marker_dict = {x : marker_wheel[i] for i, x in enumerate(default_keys)}
+
+def pretty_names(pname):
+    if pname == 'runtime':
+        return 'Runtime'
+    elif pname == 'comptime':
+        return 'Compilation time'
+    elif pname == 'overhead':
+        return 'Kernel overhead'
+    return pname
+
+#color schemes
+color_wheel = ['r', 'b', 'g', 'k']
+color_dict = {x : color_wheel[i] for i, x in enumerate(default_keys)}
+
+
+def finalize(tight=True):
+    ax = plt.gca()
+    for item in ([ax.title, ax.xaxis.label, ax.yaxis.label] +
+     ax.get_xticklabels() + ax.get_yticklabels()):
+        item.set_fontsize(tick_font_size)
+
+    if tight:
+        plt.tight_layout()

### 2017-02-15 scripts/perf_plotter.py
updates

@@ -10,7 +10,16 @@ data = data_parser.parse_data()
 reacs_as_x = False
 norm = None
 
-def plotter(plot_name='plot.pdf', **filters):
+def __compare(r, name, compare_value):
+    """
+    Specialty comparison function to account for differences
+    in runtypes
+    """
+    if name == 'vecwidth' and r.vectype == 'par':
+        return True
+    return getattr(r, name) == compare_value
+
+def plotter(plot_name='plot.pdf', show=True, **filters):
     #apply filters
     for f in filters:
         if filters[f] is None:
@@ -18,7 +27,7 @@ def plotter(plot_name='plot.pdf', **filters):
         assert any(data[x] for x in data), 'No data matching all filters'
         if f in data[data.keys()[0]][0]._asdict():
             for mech in data:
-                data[mech] = [x for x in data[mech] if getattr(x, f) == filters[f]]
+                data[mech] = [x for x in data[mech] if __compare(x, f, filters[f])]
     #compute data
     for mech in data:
         for run in data[mech]:
@@ -36,14 +45,22 @@ def plotter(plot_name='plot.pdf', **filters):
     diffs = [set([getattr(x, check) for x in plot_data]) for check in diff_check]
     #get # with more than 1 option
     diff_locs = [i for i in range(len(diffs)) if len(diffs[i]) > 1]
-    assert len(diff_locs) <= 1, "Don't know how to create graphs for more than one diff"
+    if any(diff_check[i] == 'vectype' for i in diff_locs):
+        #remove vecwidth
+        diff_locs.remove(next(i for i in diff_locs if diff_check[i] == 'vecwidth'))
+
+    assert len(diff_locs) <= 1, "Don't know how to create graphs for more than one diff: {}".format(
+        ', '.join(diff_check[i] for i in diff_locs))
     if not diff_locs:
         #regular plot
         gp.plot(plot_data, to_plot, norm=['runtime'])
     else:
         loc = diff_locs[0]
-        diffs = sorted(diffs[loc])
-        labels = [diff_check[loc] + ' ' + x for x in diffs]
+        try:
+            diffs = sorted(diffs[loc], key=lambda x: float(x))
+        except:
+            diffs = sorted(diffs[loc])
+        labels = [ps.pretty_names(diff_check[loc]).format(x) for x in diffs]
         for i, val in enumerate(diffs):
[…]

### 2017-02-15 scripts/plot_styles.py

@@ -34,12 +34,15 @@ marker_wheel = [('o', True), ('>', True), ('s', True)]
 marker_dict = {x : marker_wheel[i] for i, x in enumerate(default_keys)}
 
 def pretty_names(pname):
-    if pname == 'runtime':
-        return 'Runtime'
-    elif pname == 'comptime':
-        return 'Compilation time'
-    elif pname == 'overhead':
-        return 'Kernel overhead'
+    pname_dict = {'runtime' : 'Runtime',
+        'comptime' : 'Compilation time',
+        'overhead' : 'Kernel Construction Overhead',
+        'vecwidth' : 'Vector Width = {}',
+        'w' : 'Shallow SIMD',
+        'par' : 'SIMT'
+    }
+    if pname in pname_dict:
+        return pname_dict[pname]
     return pname
 
 #color schemes

### 2017-02-15 scripts/general_plotting.py
plot updates

@@ -12,10 +12,9 @@ legend_key = {'H2':r'H$_2$/CO',
               'IC5H11OH':r'IC$_5$H$_{11}$OH'
               }
 
-def plot(plotdata, to_plot, minx=None, miny=None, plot_std=True,
-         norm={}, return_y=False, reacs_as_x=False, labels=[], plot_ind=None
-         ):
-    """Plot performance as a function of reaction count.
+def process_data(plotdata, plot, reacs_as_x=True):
+    """
+    Process the data into an easily usable form
     """
     if reacs_as_x:
         plotdata = sorted(plotdata, key=lambda x: x.mechdata.num_reactions)
@@ -23,61 +22,69 @@ def plot(plotdata, to_plot, minx=None, miny=None, plot_std=True,
     else:
         plotdata = sorted(plotdata, key=lambda x: x.mechdata.num_species)
         x_vals = [x.mechdata.num_species for x in plotdata]
-    for i, plot in enumerate(to_plot):
-        y_vals = []
-        err_vals = []
-        for run in plotdata:
-            if plot in norm:
-                run_y = [np.array(getattr(data, plot)) / data.num_conditions for data in run.rundata]
-            else:
-                run_y = [getattr(data, plot) for data in run.rundata]
-            err_vals.append(np.std(run_y))
-            y_vals.append(np.mean(run_y))
 
-        if plot_ind is not None:
-            assert labels
-            marker, hollow = ps.marker_wheel[plot_ind]
-            color = ps.color_wheel[plot_ind]
-            name = labels[plot_ind]
+    y_vals = []
+    err_vals = []
+    for run in plotdata:
+        if plot == 'runtime':
+            run_y = [np.array(getattr(data, plot)) / data.num_conditions for data in run.rundata]
         else:
-            marker, hollow = ps.marker_dict[plot]
-            color = ps.color_dict[plot]
-            name = plot
+            run_y = [getattr(data, plot) for data in run.rundata]
+        err_vals.append(np.std(run_y))
+        y_vals.append(np.mean(run_y))
+    return x_vals, y_vals, err_vals
 
-        argdict = {'x':x_vals,
-                   'y':y_vals,
-                   'linestyle':'',
-                   'marker':marker,
-                   'label':ps.pretty_names(name)
-                   }
-        argdict['color'] = color
-        argdict['markeredgecolor'] = color
-        if not hollow:
-            argdict['markersize'] = ps.marker_style['size']
-        else:
[…]

### 2017-02-15 scripts/perf_plotter.py

@@ -3,6 +3,7 @@ import argparse
 import plot_styles as ps
 import matplotlib.pyplot as plt
 import general_plotting as gp
+import numpy as np
 
 run = data_parser.run
 rundata = data_parser.rundata
@@ -19,7 +20,7 @@ def __compare(r, name, compare_value):
         return True
     return getattr(r, name) == compare_value
 
-def plotter(plot_name='plot.pdf', show=True, **filters):
+def plotter(plot_name='', show=True, plot_reacs=True, norm=True, **filters):
     #apply filters
     for f in filters:
         if filters[f] is None:
@@ -45,7 +46,7 @@ def plotter(plot_name='plot.pdf', show=True, **filters):
     diffs = [set([getattr(x, check) for x in plot_data]) for check in diff_check]
     #get # with more than 1 option
     diff_locs = [i for i in range(len(diffs)) if len(diffs[i]) > 1]
-    if any(diff_check[i] == 'vectype' for i in diff_locs):
+    if len(diff_locs) > 1 and any(diff_check[i] == 'vectype' for i in diff_locs):
         #remove vecwidth
         diff_locs.remove(next(i for i in diff_locs if diff_check[i] == 'vecwidth'))
 
@@ -53,7 +54,8 @@ def plotter(plot_name='plot.pdf', show=True, **filters):
         ', '.join(diff_check[i] for i in diff_locs))
     if not diff_locs:
         #regular plot
-        gp.plot(plot_data, to_plot, norm=['runtime'])
+        for plot in to_plot:
+            gp.plot(plot, *gp.process_data(plot_data, plot, reacs_as_x=plot_reacs))
     else:
         loc = diff_locs[0]
         try:
@@ -61,12 +63,34 @@ def plotter(plot_name='plot.pdf', show=True, **filters):
         except:
             diffs = sorted(diffs[loc])
         labels = [ps.pretty_names(diff_check[loc]).format(x) for x in diffs]
+        #first pass - process data
+        x_vals = [None for _ in range(len(diffs))]
+        y_vals = [None for _ in range(len(diffs))]
+        z_vals = [None for _ in range(len(diffs))]
         for i, val in enumerate(diffs):
             match = [x for x in plot_data if getattr(x, diff_check[loc]) == val]
-            gp.plot(match,
-                to_plot, norm=['runtime'], labels=labels, plot_ind=i)
-    plt.ylabel(r'Runtime ($\frac{\si{\milli\second}}{\text{state}}$)')
-    plt.xlabel(r'Number of Species in Model')
[…]

### 2017-02-15 scripts/plot_styles.py

@@ -33,13 +33,33 @@ default_keys = ['runtime', 'comptime', 'overhead']
 marker_wheel = [('o', True), ('>', True), ('s', True)]
 marker_dict = {x : marker_wheel[i] for i, x in enumerate(default_keys)}
 
+class dummy_formatter(object):
+    def __init__(self, choicedict):
+        self.choices = choicedict.keys()
+        self.values = [choicedict[x] for x in self.choices]
+
+    def __hash__(self):
+        return hash(repr(self))
+
+    def format(self, val):
+        return next(self.values[i] for i, x in enumerate(self.choices)
+            if x == val)
+
 def pretty_names(pname):
     pname_dict = {'runtime' : 'Runtime',
         'comptime' : 'Compilation time',
         'overhead' : 'Kernel Construction Overhead',
         'vecwidth' : 'Vector Width = {}',
-        'w' : 'Shallow SIMD',
-        'par' : 'SIMT'
+        'vectype' : dummy_formatter({'w' : 'Shallow SIMD',
+            'par' : 'SIMT',
+            'd' : 'Deep SIMD'}),
+        'order' : dummy_formatter({'C' : 'C-order',
+            'F' : 'F-order'}),
+        'kernel' : dummy_formatter({'single' : 'Single Rate Kernel',
+            'split' : 'Split Rate Kernels'}),
+        'rates' : dummy_formatter({'fixed' : 'Fixed Rate Specialization',
+            'hybrid' : 'Hybrid Rate Specialization',
+            'full' : 'Full Rate Specialization'})
     }
     if pname in pname_dict:
         return pname_dict[pname]

### 2017-02-15 scripts/perf_plotter.py
handle 2 diffs

@@ -4,6 +4,7 @@ import plot_styles as ps
 import matplotlib.pyplot as plt
 import general_plotting as gp
 import numpy as np
+import itertools
 
 run = data_parser.run
 rundata = data_parser.rundata
@@ -46,45 +47,70 @@ def plotter(plot_name='', show=True, plot_reacs=True, norm=True, **filters):
     diffs = [set([getattr(x, check) for x in plot_data]) for check in diff_check]
     #get # with more than 1 option
     diff_locs = [i for i in range(len(diffs)) if len(diffs[i]) > 1]
-    if len(diff_locs) > 1 and any(diff_check[i] == 'vectype' for i in diff_locs):
-        #remove vecwidth
-        diff_locs.remove(next(i for i in diff_locs if diff_check[i] == 'vecwidth'))
-
-    assert len(diff_locs) <= 1, "Don't know how to create graphs for more than one diff: {}".format(
-        ', '.join(diff_check[i] for i in diff_locs))
+    diffs = [x for x in diffs if len(x) > 1]
+    if len(diff_locs) > 2:
+        raise NotImplementedError
     if not diff_locs:
         #regular plot
         for plot in to_plot:
             gp.plot(plot, *gp.process_data(plot_data, plot, reacs_as_x=plot_reacs))
     else:
-        loc = diff_locs[0]
+        #create map dict
+        loc_map = {}
+        for i, diff in enumerate(diffs):
+            for subdiff in diff:
+                loc_map[subdiff] = diff_locs[i]
+
+        #sort
         try:
-            diffs = sorted(diffs[loc], key=lambda x: float(x))
+            diffs = [sorted(diff, key=lambda x: float(x)) for diff in diffs]
         except:
-            diffs = sorted(diffs[loc])
-        labels = [ps.pretty_names(diff_check[loc]).format(x) for x in diffs]
+            diffs = [sorted(diff) for diff in diffs]
+
         #first pass - process data
-        x_vals = [None for _ in range(len(diffs))]
-        y_vals = [None for _ in range(len(diffs))]
-        z_vals = [None for _ in range(len(diffs))]
-        for i, val in enumerate(diffs):
-            match = [x for x in plot_data if getattr(x, diff_check[loc]) == val]
-            x_vals[i], y_vals[i], z_vals[i] = gp.process_data(match, 'runtime', reacs_as_x=plot_reacs)
+        x_vals = []
+        y_vals = []
+        z_vals = []
[…]

### 2017-02-15 scripts/plot_styles.py

@@ -30,7 +30,8 @@ clear_marker_style = {
 }
 
 default_keys = ['runtime', 'comptime', 'overhead']
-marker_wheel = [('o', True), ('>', True), ('s', True)]
+marker_wheel = [('o', True), ('>', True), ('s', True),
+    ('D', True)]
 marker_dict = {x : marker_wheel[i] for i, x in enumerate(default_keys)}
 
 class dummy_formatter(object):

### 2017-02-16 scripts/general_plotting.py
fix core plotter, updated perf

@@ -12,13 +12,17 @@ legend_key = {'H2':r'H$_2$/CO',
               'IC5H11OH':r'IC$_5$H$_{11}$OH'
               }
 
-def process_data(plotdata, plot, reacs_as_x=True):
+def process_data(plotdata, plot, reacs_as_x=True,
+    plot_cores=False):
     """
     Process the data into an easily usable form
     """
     if reacs_as_x:
         plotdata = sorted(plotdata, key=lambda x: x.mechdata.num_reactions)
         x_vals = [x.mechdata.num_reactions for x in plotdata]
+    elif plot_cores:
+        plotdata = sorted(plotdata, key=lambda x: float(x.cores))
+        x_vals = [float(x.cores) for x in plotdata]
     else:
         plotdata = sorted(plotdata, key=lambda x: x.mechdata.num_species)
         x_vals = [x.mechdata.num_species for x in plotdata]

### 2017-02-16 scripts/perf_plotter.py

@@ -1,3 +1,5 @@
+from __future__ import division
+
 import data_parser
 import argparse
 import plot_styles as ps
@@ -5,6 +7,7 @@ import matplotlib.pyplot as plt
 import general_plotting as gp
 import numpy as np
 import itertools
+import functools
 
 run = data_parser.run
 rundata = data_parser.rundata
@@ -12,14 +15,26 @@ data = data_parser.parse_data()
 reacs_as_x = False
 norm = None
 
-def __compare(r, name, compare_value):
+#from http://stackoverflow.com/a/31174427
+sentinel = object()
+def rgetattr(obj, attr, default=sentinel):
+    if default is sentinel:
+        _getattr = getattr
+    else:
+        def _getattr(obj, name):
+            return getattr(obj, name, default)
+    return functools.reduce(_getattr, [obj]+attr.split('.'))
+
+def __compare(r, name, compare_value, plot_cores=False):
     """
     Specialty comparison function to account for differences
     in runtypes
     """
     if name == 'vecwidth' and r.vectype == 'par':
         return True
-    return getattr(r, name) == compare_value
+    if plot_cores and name == 'cores':
+        return True
+    return rgetattr(r, name) == compare_value
 
 def plotter(plot_name='', show=True, plot_reacs=True, norm=True, **filters):
     #apply filters
@@ -48,6 +63,19 @@ def plotter(plot_name='', show=True, plot_reacs=True, norm=True, **filters):
     #get # with more than 1 option
     diff_locs = [i for i in range(len(diffs)) if len(diffs[i]) > 1]
     diffs = [x for x in diffs if len(x) > 1]
+
+    plot_cores = False
+    if 'cores' in [diff_check[loc] for loc in diff_locs]:
+        #can only process one
+        assert len(diff_locs) == 1
+        diffs = [sorted(data.keys())]
+        diff_locs = [-1]
+        diff_check.append('mechdata.mech')
+        plot_reacs = False
+        plot_cores = True
+
+    #delete diff
+
     if len(diff_locs) > 2:
         raise NotImplementedError
     if not diff_locs:
@@ -77,27 +105,26 @@ def plotter(plot_name='', show=True, plot_reacs=True, norm=True, **filters):
         if len(diffs) == 1:
             for val in [subdiff for diff in diffs for subdiff in diff]:
                 check = diff_check[loc_map[val]]
-                match = [x for x in plot_data if __compare(x, check, val)]
[…]

### 2017-02-16 scripts/plot_styles.py

@@ -9,7 +9,7 @@ import os.path
 plt.rc('text', usetex=True)
 plt.rc('font', family='serif')
 plt.rc('text.latex',
-    preamble=r'\usepackage{amsmath},\usepackage{siunitx}')
+    preamble=r'\usepackage{amsmath},\usepackage{siunitx},\usepackage[version=3]{mhchem}')
 plt.rc('font', family='serif')
 
 legend_style = {'loc':0,
@@ -60,7 +60,8 @@ def pretty_names(pname):
             'split' : 'Split Rate Kernels'}),
         'rates' : dummy_formatter({'fixed' : 'Fixed Rate Specialization',
             'hybrid' : 'Hybrid Rate Specialization',
-            'full' : 'Full Rate Specialization'})
+            'full' : 'Full Rate Specialization'}),
+        'mechdata.mech' : r'\ce{{{}}}' #used in core plotting
     }
     if pname in pname_dict:
         return pname_dict[pname]

### 2017-02-16 scripts/perf_plotter.py
fix vectype

@@ -74,7 +74,12 @@ def plotter(plot_name='', show=True, plot_reacs=True, norm=True, **filters):
         plot_reacs = False
         plot_cores = True
 
-    #delete diff
+    #delete diff for vecwidth / par thing
+    if 'vectype' in [diff_check[loc] for loc in diff_locs]:
+        ind = next(i for i, loc in enumerate(diff_locs)
+            if diff_check[loc] == 'vecwidth')
+        diffs.pop(ind)
+        diff_locs.pop(ind)
 
     if len(diff_locs) > 2:
         raise NotImplementedError

### 2017-02-16 scripts/perf_plotter.py
fix

@@ -76,10 +76,11 @@ def plotter(plot_name='', show=True, plot_reacs=True, norm=True, **filters):
 
     #delete diff for vecwidth / par thing
     if 'vectype' in [diff_check[loc] for loc in diff_locs]:
-        ind = next(i for i, loc in enumerate(diff_locs)
-            if diff_check[loc] == 'vecwidth')
-        diffs.pop(ind)
-        diff_locs.pop(ind)
+        ind = next((i for i, loc in enumerate(diff_locs)
+            if diff_check[loc] == 'vecwidth'), None)
+        if ind is not None:
+            diffs.pop(ind)
+            diff_locs.pop(ind)
 
     if len(diff_locs) > 2:
         raise NotImplementedError

### 2017-02-16 paper.tex
update validation disc

@@
 \begin{equation}
-\frac{\text{d}[T]}{\text{d}t} = -\frac{\sum_{k}^{N_S} H^{\degree}_k \dot{\omega}_k}{\sum_{k}^{N_S} [C_k] c_{p_k}^{\degree}}
+\frac{\text{d}[T]}{\text{d}t} = -\frac{\sum_{k}^{N_S} H^{\degree}_k \frac{\text{d}[C_k]}{\text{d}t}}{\sum_{k}^{N_S} [C_k] c_{p_k}^{\degree}}
 \end{equation}
@@
 \subsection{Validation}
-The reaction rates of progress, species and temperature rates in this study are validated by comparison to Cantera~\cite{Cantera}, however due to the intricacies of floating point arithmetic care must be taken when comparing results.
+The reaction rates of progress (ROP), species and temperature rates in this study are validated by comparison with Cantera~\cite{Cantera}, however special care must be taken due floating point arithmetic issues
 
-In obtaining the net rate of progress of reaction $i$ from the forward and reverse rates of progress, $R_{i} = R_{i}^{\prime} - R_{i}^{\prime\prime}$, a base-10 bound on the loss of precision can be found as~\cite{goldberg1991every}:
+In computing the net rate of progress of reaction $i$ from the forward and reverse ROP: $R_{i} = R_{i}^{\prime} - R_{i}^{\prime\prime}$, precision can easily be lost as the net rate of progress may be---particularly near chemical equilibrium---many orders of magnitude smaller than the forward or reverse rates.
+A base-ten bound on loss of precision~\cite{goldberg1991every} can be found as:
 \begin{equation}
-2^{-p_i} \le \left\lvert 1 - \frac{R_{i}^{\prime}}{R_{i}^{\prime\prime}} \right\rvert \le 2^{-q_i}
+\text{P}_{\text{min},i} = 2^{-p_i} \le \left\lvert 1 - \frac{R_{i}^{\prime}}{R_{i}^{\prime\prime}} \right\rvert \le 2^{-q_i} = \text{P}_{\text{max},i}
 \label{e:bounds}
 \end{equation}
-with $q_i \le p_i$, while $1 - \frac{R_{i}^{\prime}}{R_{i}^{\prime\prime}}$ can be computed with accurately via use of higher-precision numbers (in this case, via use of Python's Decimal class).
-
-Plugging the forward and reverse rate constants from Cantera into Eq.~\eqref{e:bounds} we can obtain a reasonable a-priori precision bounds $\text{P}_{\text{min},i} = 2^{-p_i}$, $\text{P}_{\text{max},i} = 2^{-q_i}$.
-Finally, when determining the error in the net rates of progress between this code and Cantera, the range of percent error attributable to precision loss, $\text{P}_{\text{err},i}$ can be found as:
+with $q_i \le p_i$.
+To obtain an estimate of the a-priori precision bounds ($\text{P}_{\text{min},i}$ and $\text{P}_{\text{max},i}$), the forward and reverse ROP from Cantera are substituted into Eq.~\eqref{e:bounds}.
+Additionally a higher-precision number format (in this case, the Python Decimal class) is used in this computation to maintain accuracy.
+Finally, when determining the error in the net ROP between this code and Cantera, the bounds on the percent error attributable to precision loss, $\text{P}_{\text{err},i}$ can be estimated as:
 \begin{equation}
@@
 where the \text{CT} subscript indicates values from Cantera.
-The error of a quantity $X_{i,j}$ for state $j$, and index $i$ was determined as:
+
+For a more direct comparison, a relative error norm of a quantity $X_{i,j}$ over all states $j$, and indices $i$ was computed using the $L^{\infty}$ norm:
 \begin{equation}
- E_{X_{i}} = \left\lVert \left\lVert \frac{\left\lvert X_{i,j,\text{CT}} - X_{i,j}\right\rvert}{\num{e-10} + \num{e-6} * \left\lvert X_{i,j,\text{CT}} \right\rvert} \right\rVert_{j,\infty} \right\rVert_{i, \infty}
+ E_{X_{i}} = \left\lVert \frac{\left\lvert X_{i,j,\text{CT}} - X_{i,j}\right\rvert}{\num{e-10} + \num{e-6} * \left\lvert X_{i,j,\text{CT}} \right\rvert} \right\rVert_{i,j,\infty}
 \label{e:rel_err}
 \end{equation}
-Finally, to estimate the effect of loss of precision on the net rates of progress, we let $P_{\text{min}}^{*} = P_{i,\text{min}}$ for the $i,j$ value that results in $E_{R_{i}}$.
+and the lower precision bound for the net rate of progress error norm is set to $P_{\text{min}}^{*} = P_{i,\text{min}}$ for the $i, j$ value that results in the maximum error $E_{R_i}$.
 
@@
 \toprule
-\multicolumn{1}{c}{Model} & \multicolumn{1}{c}{$E_{R_{i}^{\prime}}$} & \multicolumn{1}{c}{$E_{R_{i}^{\prime\prime}}$} &\multicolumn{1}{c}{$E_{R_{i}}$} & \multicolumn{1}{c}{$\text{P}_{\text{min}}^{*}$} & \multicolumn{1}{c}{$E_{\frac{\text{d}[C_k]}{\text{d}t}}$} & \multicolumn{1}{c}{$E_{\frac{\text{d}[T]}{\text{d}t}}$} \\
+\multicolumn{1}{c}{Model} & \multicolumn{1}{c}{$E_{R_{i}^{\prime}}$} & \multicolumn{1}{c}{$E_{R_{i}^{\prime\prime}}$} &\multicolumn{1}{c}{$E_{R_{i}}$} & \multicolumn{1}{c}{$\text{P}_{\text{min}}^{*}\si{\percent}$} & \multicolumn{1}{c}{$E_{\frac{\text{d}[C_k]}{\text{d}t}}$} & \multicolumn{1}{c}{$E_{\frac{\text{d}[T]}{\text{d}t}}$} \\
 \midrule
@@
 
-In Table~\ref{T:error}, we see the results of this code as compared to Cantera on a library of partially stirred reaction conditions described in our previous works~\cite{CurtisGPU:2017,Niemeyer:2016aa}
+In Table~\ref{T:error}, we see the results of this code as compared to Cantera on a library of partially stirred reaction conditions described in our previous works~\cite{CurtisGPU:2017,Niemeyer:2016aa}.
+Very close agreement is observed for the forward and reverse ROP for all models, however the net rate of progress is \numrange{7}{8} orders of magnitude larger.
+We see here that the minimum estimated loss of precision is \SIrange{4}{40}{$\times$} the error in net ROP, explaining this jump.
+The species rate error norm is similar in magnitude to that of the net ROP, however the temperature rate error again increases as it depends directly on the species rates.
+We note that the above discussion does not imply that computation of the net ROP will cause large error in chemical kinetic integration---either in this code or Cantera---as this loss of precision only occurs when the net ROP are many orders of magnitude smaller than the forward or reverse rates, implying the reaction is in near-equilibrium.
 

### 2017-02-16 paper.tex
revise vectorization discusson

@@
 For a chemical kinetic model with species $k$ for the $j$-th thermo-chemical state to be evaluated, $[C_k]_j$ there are two choices for data-storage in memory.
-In the ``C'' (row-major) format, species concentrations for $j$-th thermo-chemical state are stored sequentially in memory, i.e.: $[C_0]_0\text{,}\allowbreak C[1]_0\ldots\allowbreak [C_{N_S}]_0\text{,}\allowbreak [C_0]_1\allowbreak\ldots\allowbreak [C_{N_S}]_N$, where $N$ is the total number of thermo-chemical conditions.
-Alternatively, in the ``F'' (Fortan column-major) format species concentrations a single species for all thermo-chemical states are sequential, i.e.: $[C_0]_0\text{,}\allowbreak[C_0]_1\ldots\allowbreak[C_0]_N\text{,}\allowbreak[C_1]_0\ldots\allowbreak[C_{N_S}]_N$
+In the ``C'' (C row-major) format, species concentrations for $j$-th thermo-chemical state are stored sequentially in memory.
+Alternatively, in the ``F'' (Fortan column-major) format species concentrations a single species for over all thermo-chemical states are adjacent in memory.
 In Sec.~\ref{S:results} the effect of this ordering will be shown to have a large performance impact on the program runtime.
 
-In this work, a \textit{shallow}-SIMD vectorization has been implemented, entailing each SIMD-lane evaluating the species rates for an thermo-chemical state independently.
+In this work, a \textit{shallow}-SIMD vectorization has been implemented, entailing each SIMD-lane evaluating chemical source terms for an thermo-chemical state independently.
 In future efforts, a so called \textit{deep}-SIMD vectorization will be pursued allowing vector-units to cooperate to evaluate species rates for a single thermo-chemical state.
 A deep-vectorization may result in SIMD-\textit{waste} in this context---similar to thread-divergence in SIMT-processing---caused by different SIMD-lanes executing different instructions (e.g.
-from differing for-loop bounds), however past studies demonstrate the merits of deep-SIMT vectorization on GPUs~\cite{Sewerin20151375}.
+from differing for-loop bounds).
+However a deep-vectorization will likely have higher cache hit-rates than a shallow-vectorization using the ``C'' memory layout as successive SIMD-lanes would access directly memory adjacent locations.
+Additionally, past studies demonstrate the merits of deep-SIMT vectorization on GPUs~\cite{Sewerin20151375}.
 
@@
 Error statistics are based on the infinity-norm of the relative error detailed in Eq.~\eqref{e:rel_err} for each quantity.
+\todo[inline]{i-pentanol validation still running}
 }

### 2017-02-16 scripts/general_plotting.py
add ability to specify markers / labels / legends; add simd plotter'

@@ -40,12 +40,14 @@ def process_data(plotdata, plot, reacs_as_x=True,
 
 
 def plot(plot, x_vals, y_vals, err_vals, minx=None, miny=None, plot_std=True,
-    return_y=False, labels=[], plot_ind=None
-         ):
+        return_y=False, labels=[], plot_ind=None, marker_func=None):
     """Plot performance as a function of reaction count.
     """
 
-    if plot_ind is not None:
+    if marker_func is not None:
+        name = labels[plot_ind]
+        marker, hollow, color = marker_func(name)
+    elif plot_ind is not None:
         assert labels
         marker, hollow = ps.marker_wheel[plot_ind]
         color = ps.color_wheel[plot_ind]

### 2017-02-16 scripts/perf_plotter.py

@@ -36,7 +36,11 @@ def __compare(r, name, compare_value, plot_cores=False):
         return True
     return rgetattr(r, name) == compare_value
 
-def plotter(plot_name='', show=True, plot_reacs=True, norm=True, **filters):
+def plotter(plot_name='', show=True, plot_reacs=True, norm=True,
+        legend_handler=None, marker_func=None, **filters):
+    #create fig, ax
+    fig = plt.figure()
+    ax = plt.subplot(1,1,1)
     #apply filters
     for f in filters:
         if filters[f] is None:
@@ -51,9 +55,9 @@ def plotter(plot_name='', show=True, plot_reacs=True, norm=True, **filters):
             rund = run.rundata
     #now plot data
     to_plot = ['runtime']
-    if filters['plot_compilation']:
+    if filters.pop('plot_compilation', False):
         to_plot.append('comptime')
-    if filters['plot_overhead']:
+    if filters.pop('plot_overhead', False):
         to_plot.append('overhead')
     #get data
     plot_data = [x for mech in data for x in data[mech]]
@@ -107,6 +111,28 @@ def plotter(plot_name='', show=True, plot_reacs=True, norm=True, **filters):
         z_vals = []
         labels = []
 
+        def __get_compound_names(val1, val2):
+            c1 = diff_check[loc_map[val1]]
+            c2 = diff_check[loc_map[val2]]
+
+            #generic name ordering
+            ordering = ['vectype', 'order']
+
+            #sort by order
+            check_vals = [None for _ in diff_check]
+            check_vals[loc_map[val1]] = c1
+            check_vals[loc_map[val2]] = c2
+            check_vals = sorted(check_vals,
+                key=lambda x: 100 if x not in ordering else ordering.index(x))
+
+            #remove none
+            check_vals = [x for x in check_vals if x is not None]
+
+            #and return str
+            return ' - '.join(ps.pretty_names(check).format(val1 if check == c1 else val2)
+                for check in check_vals)
+
+
         #handle 2 diffs
         if len(diffs) == 1:
             for val in [subdiff for diff in diffs for subdiff in diff]:
@@ -123,7 +149,7 @@ def plotter(plot_name='', show=True, plot_reacs=True, norm=True, **filters):
                 match = [x for x in plot_data if __compare(x, diff_check[loc_map[val1]], val1, plot_cores=plot_cores)
[…]

### 2017-02-16 scripts/plot_styles.py

@@ -46,7 +46,7 @@ class dummy_formatter(object):
         return next(self.values[i] for i, x in enumerate(self.choices)
             if x == val)
 
-def pretty_names(pname):
+def pretty_names(pname, short=False):
     pname_dict = {'runtime' : 'Runtime',
         'comptime' : 'Compilation time',
         'overhead' : 'Kernel Construction Overhead',

### 2017-02-16 scripts/simd_and_order_comp.py

@@ -1,0 +1,46 @@
+import perf_plotter as pp
+import plot_styles as ps
+import matplotlib.patches as mpatches
+import matplotlib.lines as mlines
+
+SIMD_color = ps.color_wheel[0]
+SIMT_color = ps.color_wheel[1]
+SIMT_sym = ps.marker_wheel[0][0]
+SIMD_sym = ps.marker_wheel[1][0]
+C_filled = True
+F_filled = False
+
+def marker_func(name):
+    retval = []
+    if 'SIMT' in name:
+        retval.append(SIMT_sym)
+    else:
+        retval.append(SIMD_sym)
+    if 'C-order' in name:
+        retval.append(C_filled)
+    else:
+        retval.append(F_filled)
+    if 'SIMT' in name:
+        retval.append(SIMT_color)
+    else:
+        retval.append(SIMD_color)
+    return tuple(retval)
+
+labels = ['SIMT', 'SIMD', 'C-order', 'F-order']
+handles = [mlines.Line2D([], [], color=SIMT_color, marker=SIMT_sym, linestyle='',
+                          markersize=ps.marker_style['size'], label='SIMT',
+                          markerfacecoloralt='none',
+                          fillstyle='left'),
+           mlines.Line2D([], [], color=SIMD_color, marker=SIMD_sym, linestyle='',
+                          markersize=ps.marker_style['size'], label='SIMD',
+                          markerfacecoloralt='none',
+                          fillstyle='left'),
+           mlines.Line2D([], [], linestyle='', marker=(4, 0, 45), markersize=ps.marker_style['size'],
+                          label='C-order', markerfacecolor='k', markeredgecolor='k'),
+           mlines.Line2D([], [], linestyle='', marker=(4, 0, 45), markersize=ps.marker_style['size'],
+                          label='F-order', markerfacecolor='none', markeredgecolor='k')]
+legend = (handles, labels)
+
+fig, ax = pp.plotter(show=True,
+    cores='1', vecwidth='8', rates='hybrid', kernel='split',
+    legend_handler=legend, marker_func=marker_func)

### 2017-02-16 paper.tex
add specific fig plotters, initial versions of figs and update bib

@@
 \usepackage{booktabs,multicol} %better tables
+\usepackage{subcaption} %subfigs
 \sisetup{group-separator={,},
@@
 
-\begin{figure}[ht]
+\begin{figure}[htb]
 	\centering
@@
 a CPU core) contains a vector unit with several lanes (L0, L1, etc.).
-The vector unit executes a single instruction concurrently on multiple data.
-Adapted from\cite{Simdfig:2016}}
+The vector unit executes a single instruction concurrently on multiple data.}
 	\label{F:SIMD}
@@
 
-\begin{table}[ht]
+\begin{table}[htb]
 \sisetup{retain-zero-exponent=true}
@@
 
-In Table~\ref{T:error}, we see the results of this code as compared to Cantera on a library of partially stirred reaction conditions described in our previous works~\cite{CurtisGPU:2017,Niemeyer:2016aa}.
+In Table~\ref{T:error}, we see the results of this code as compared to Cantera on a library of partially stirred reaction conditions (PaSR) described in our previous works~\cite{CurtisGPU:2017,Niemeyer:2016aa}.
 Very close agreement is observed for the forward and reverse ROP for all models, however the net rate of progress is \numrange{7}{8} orders of magnitude larger.
@@
 \label{S:results}
+The performance studies in this work were run on four ten-core \SI{2.2}{\giga\hertz} Intel Xeon E5-4640 v2 CPUs with \SI{20}{\mega\byte} of L3 cache memory, installed on an Ace Powerworks PW8027R-TRF+ with a Supermicro X9QR7-TF+/X9QRi-F+ baseboard.
+Runtimes in each case were averaged over ten runs, each using the same set of PaSR conditions utilized in validation.
+All code was compiled with \texttt{gcc 4.8.5} and ``\texttt{-O3}'' optimization, running on \texttt{v16.1.1} of the Intel OpenCL runtime supporting OpenCL \texttt{v1.2}.
+
+\begin{figure}[htb]
+  \centering
+  \begin{subfigure}[t]{0.5\textwidth}
+    \includegraphics[width=\textwidth]{SIMD_SIMT_Order_comparison.pdf}
+    \caption{}
+    \label{F:SIMDComp}
+  \end{subfigure}
+  ~
+
+\end{figure}
 

### 2017-02-16 scripts/general_plotting.py

@@ -39,7 +39,7 @@ def process_data(plotdata, plot, reacs_as_x=True,
     return x_vals, y_vals, err_vals
 
 
-def plot(plot, x_vals, y_vals, err_vals, minx=None, miny=None, plot_std=True,
+def plot(plot, x_vals, y_vals, err_vals, minx=None, miny=None, maxx=None, maxy=None, plot_std=True,
         return_y=False, labels=[], plot_ind=None, marker_func=None):
     """Plot performance as a function of reaction count.
     """
@@ -80,18 +80,17 @@ def plot(plot, x_vals, y_vals, err_vals, minx=None, miny=None, plot_std=True,
 
     this_minx = np.min(x_vals)
     this_miny = np.min(y_vals)
-    minx = (this_minx if minx is None
-            else this_minx if this_minx < minx
-            else minx
-            )
-    miny = (this_miny if miny is None
-            else this_miny if this_miny < miny
-            else miny
-            )
-    retval = minx, miny
-    if return_y:
-        retval = (retval, y_vals, err_vals)
-    return retval
+    def __get_min(test, inval):
+        return test if inval is None else (
+            test if test < inval else inval)
+    def __get_max(test, inval):
+        return test if inval is None else (
+            test if test > inval else inval)
+    minx = __get_min(np.min(x_vals), minx)
+    miny = __get_min(np.min(y_vals), miny)
+    maxx = __get_min(np.max(x_vals), maxx)
+    maxy = __get_min(np.max(y_vals), maxy)
+    return minx, miny, maxx, maxy
 
 
 def plot_scaling(plotdata, markerlist, colorlist, minx=None, miny=None,

### 2017-02-16 scripts/perf_plotter.py

@@ -8,10 +8,11 @@ import general_plotting as gp
 import numpy as np
 import itertools
 import functools
+import copy
 
 run = data_parser.run
 rundata = data_parser.rundata
-data = data_parser.parse_data()
+data_clean = data_parser.parse_data()
 reacs_as_x = False
 norm = None
 
@@ -37,7 +38,9 @@ def __compare(r, name, compare_value, plot_cores=False):
     return rgetattr(r, name) == compare_value
 
 def plotter(plot_name='', show=True, plot_reacs=True, norm=True,
-        legend_handler=None, marker_func=None, **filters):
+        legend_handler=None, marker_func=None,
+        minx=None, miny=None, maxx=None, maxy=None, **filters):
+    data = copy.copy(data_clean)
     #create fig, ax
     fig = plt.figure()
     ax = plt.subplot(1,1,1)
@@ -71,12 +74,13 @@ def plotter(plot_name='', show=True, plot_reacs=True, norm=True,
     plot_cores = False
     if 'cores' in [diff_check[loc] for loc in diff_locs]:
         #can only process one
-        assert len(diff_locs) == 1
-        diffs = [sorted(data.keys())]
-        diff_locs = [-1]
-        diff_check.append('mechdata.mech')
-        plot_reacs = False
-        plot_cores = True
+        plot_cores = len(diff_locs) == 1
+        if plot_cores:
+            diffs = [sorted(data.keys())]
+            diff_locs = [-1]
+            diff_check.append('mechdata.mech')
+            plot_reacs = False
+            plot_cores = True
 
     #delete diff for vecwidth / par thing
     if 'vectype' in [diff_check[loc] for loc in diff_locs]:
@@ -143,7 +147,7 @@ def plotter(plot_name='', show=True, plot_reacs=True, norm=True,
                     x, y, z = gp.process_data(match, 'runtime', reacs_as_x=plot_reacs, plot_cores=plot_cores)
                     x_vals.append(x); y_vals.append(y); z_vals.append(z)
         else:
-            iterator = [zip(x,diffs[1]) for x in itertools.permutations(diffs[0],len(diffs[1]))]
+            iterator = [zip(x,diffs[1]) for x in itertools.permutations(diffs[0],len(diffs[0]))]
             iterator = [subiter for i in iterator for subiter in i]
             for val1, val2 in iterator:
[…]

### 2017-02-16 scripts/simd_and_order_comp.py

@@ -7,8 +7,8 @@ SIMD_color = ps.color_wheel[0]
 SIMT_color = ps.color_wheel[1]
 SIMT_sym = ps.marker_wheel[0][0]
 SIMD_sym = ps.marker_wheel[1][0]
-C_filled = True
-F_filled = False
+C_hollow = False
+F_hollow = True
 
 def marker_func(name):
     retval = []
@@ -17,9 +17,9 @@ def marker_func(name):
     else:
         retval.append(SIMD_sym)
     if 'C-order' in name:
-        retval.append(C_filled)
+        retval.append(C_hollow)
     else:
-        retval.append(F_filled)
+        retval.append(F_hollow)
     if 'SIMT' in name:
         retval.append(SIMT_color)
     else:
@@ -41,6 +41,6 @@ handles = [mlines.Line2D([], [], color=SIMT_color, marker=SIMT_sym, linestyle=''
                           label='F-order', markerfacecolor='none', markeredgecolor='k')]
 legend = (handles, labels)
 
-fig, ax = pp.plotter(show=True,
+pp.plotter(show=True,
     cores='1', vecwidth='8', rates='hybrid', kernel='split',
     legend_handler=legend, marker_func=marker_func)

### 2017-02-16 paper.tex
update data ordering discussion

@@
 \subsection{Data Ordering and Vectorization Patterns}
-For a chemical kinetic model with species $k$ for the $j$-th thermo-chemical state to be evaluated, $[C_k]_j$ there are two choices for data-storage in memory.
-In the ``C'' (C row-major) format, species concentrations for $j$-th thermo-chemical state are stored sequentially in memory.
-Alternatively, in the ``F'' (Fortan column-major) format species concentrations a single species for over all thermo-chemical states are adjacent in memory.
-In Sec.~\ref{S:results} the effect of this ordering will be shown to have a large performance impact on the program runtime.
+\label{S:data}
+When storing arrays for a chemical kinetic model, e.g. the concentration of species $k$ for the $j$-th of thermo-chemical state $[C_k]_j$ there are two data-ordering choices.
+In the ``F'' (Fortran column-major) format, the concentrations of a single species $k$ over all $N$ thermo-chemical states are adjacent: $[C_k]_j, [C_k]_{j + 1}\ldots[C_k]_{N}$;
+conversely in the ``C'' (C row-major) format, the species concentrations for the $j$-th thermo-chemical state: $[C_k]_j, [C_{k+1}]_j\ldots$ are stored sequentially in memory;
+This ordering---along with the device (CPU, GPU, etc.) in question---have a large effect on the performance of SIMD\slash SIMT-vectorized algorithms.
 
-In this work, a \textit{shallow}-SIMD vectorization has been implemented, entailing each SIMD-lane evaluating chemical source terms for an thermo-chemical state independently.
-In future efforts, a so called \textit{deep}-SIMD vectorization will be pursued allowing vector-units to cooperate to evaluate species rates for a single thermo-chemical state.
-A deep-vectorization may result in SIMD-\textit{waste} in this context---similar to thread-divergence in SIMT-processing---caused by different SIMD-lanes executing different instructions (e.g.
+In this work a \textit{shallow}-SIMD vectorization, in which each SIMD-lane evaluates chemical source terms for a different thermo-chemical state, has been implemented.
+If the data is stored in ``F''-order, the SIMD-lanes will access sequential locations in memory however the next species' concentration---for example---will be $N$ memory locations away, potentially resulting in more cache-misses on the CPU.
+In future efforts, a so called \textit{deep}-SIMD vectorization will be pursued, in which a vector-unit cooperates to evaluate chemical source terms for a single thermo-chemical state.
+In this context, a deep-vectorization may result in SIMD-\textit{waste}---similar to thread-divergence in GPU-processing---caused by different SIMD-lanes executing different instructions (e.g.
 from differing for-loop bounds).
-However a deep-vectorization will likely have higher cache hit-rates than a shallow-vectorization using the ``C'' memory layout as successive SIMD-lanes would access directly memory adjacent locations.
-Additionally, past studies demonstrate the merits of deep-SIMT vectorization on GPUs~\cite{Sewerin20151375}.
+However, SIMD-lanes in a ``C''-ordered deep-vectorization access sequential memory locations and further the data locality of e.g., species concentrations, will likely result in improved CPU cache-hit rates.
+Past studies additionally demonstrate promising results for deep-SIMT vectorization on GPUs~\cite{Sewerin20151375}.
+In Sec.~\ref{S:results} the effect of this ordering on the shallow-SIMD vectorization performance will be studied.
 

### 2017-02-17 paper.tex
move data access / vec patterns discussion up to intro to save space in discussion of past work; update plots

@@
   \centering
-	\includegraphics[width=0.33\linewidth]{SIMD.pdf}
+  \begin{subfigure}[t]{0.54\linewidth}
+      \includegraphics[width=\textwidth]{SIMD.pdf}
       \caption{Schematic of SIMD processing.
@@
       \label{F:SIMD}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.43\linewidth}
+      \includegraphics[width=\textwidth]{mem_access.pdf}
+      \caption{Representation of an 2-D array with dimension $N_1 \times N_2$ in CPU memory, with linear index marked.}
+      \label{F:mem}
+  \end{subfigure}
 \end{figure}
@@
 
-Several recent works, e.g.
-\cite{CurtisGPU:2017,Sewerin20151375} have investigated SIMT-based chemical kinetic integration on GPUs, however SIMD-based chemical kinetics have been less studied in comparison.
-\textcite{stone2016} implemented a linearly-implicit fourth-order stiff Rosenbrock solver in the OpenCL for various platforms including the CPU, GPU and the MIC; the Rosenbrock solver performance was compared to that of a non-stiff, fourth-order explicit Runge–Kutta–Fehlberg integrator.
-Two parallelization models were investigated in OpenCL, the so called \textit{shallow}-SIMD vectorization, where each SIMD-lane solved a single chemical kinetic initial value problem (IVP), and a SIMT-vectorization where each thread solved a single IVP.
-Shallow-SIMD vectorization improved the integrator performance over the OpenMP baseline by \SIrange{2.5}{2.8}{$\times$} on the CPU and \SIrange{4.7}{4.9}{$\times$} on the MIC, while the GPU performance while the GPU was \SIrange{1.4}{1.6}{$\times$} slower than the OpenMP baseline due to thread-divergence concerns.
-\textcite{kroshko2013efficient} implemented a shallow-SIMD -vectorized third order Rosenbrock solver (RODAS) integrator for atmospheric chemistry on a Cell Broadband Engine---a specially designed vector processor---finding a speedup of \SI{1.89}{$\times$} over a serial version of the same code, a \SI{94}{\percent} parallel efficiency.
+\subsection{Data ordering and vectorization patterns}
+\label{S:data}
+When storing arrays for a chemical kinetic model, e.g. the concentration of species $k$ for the $j$-th of thermo-chemical state $[C_k]_j$ there are two data-ordering choices.
+In the ``F'' (Fortran column-major) format, the concentrations of a single species $k$ over all $N$ thermo-chemical states are adjacent.
+In Fig.~\ref{F:mem} this would correspond to storing $[C_0]_0$ in index \num{0}, $[C_0]_1$ in index \num{1} and so on, with $N_1 = N$, and $N_2 = N_s$, the number of species in the model.
+In the ``C'' (C row-major) format, the storage is flipped with $[C_0]_0$ in index \num{0}, $[C_1]_0$ in index \num{1} etc., with $N_1 = N_s$ and $N_2 = N$.
+This ordering---along with the device (CPU, GPU, etc.) in question---have a large effect on the performance of SIMD\slash SIMT-vectorized algorithms.
 
-This work will study the performance of chemical source term evaluation of automatically generated shallow-SIMD-vectorized codes for a wide range chemical kinetic models on the CPU.
-The performance will be compared to a baseline SIMT-vectorized codes to determine the effective SIMD speedup.
+In a \textit{shallow}-SIMD vectorization, each SIMD-lane evaluates chemical source terms for a different thermo-chemical state, has been implemented.
+If the data is stored in ``F''-order, the SIMD-lanes accessing the concentration of species $k$ for different states will load sequential memory locations in Fig.~\ref{F:mem} in memory however the $k+1$'th species concentration will be $N$ memory locations away, increasing the likelihood of cache-misses on the CPU.
+In a \textit{deep}-SIMD vectorization, a vector-unit cooperates to evaluate chemical source terms for a single thermo-chemical state, thus SIMD-lanes loading species concentrations $k\dots k+V$ (for a vector width of $V$) will access sequential memory locations if the data is stored in ``C''-order.
+Further, in the ``C''-ordering the furthest difference between any two concentrations in a single thermo-chemical state is at most $N_s \ll N$ (in most cases), greatly improving the data-locality and chance of a cache-hit on the CPU.
+However, a deep-vectorization may result in SIMD-\textit{waste}---similar to thread-divergence in GPU-processing---caused by different SIMD-lanes executing different instructions (e.g.
+from differing for-loop bounds).
+In Sec.~\ref{S:results} the effect of this ordering on the shallow-SIMD vectorization performance will be studied.
+
+\subsection{Previous works and goals of this tudy}
+Several recent works, e.g. \cite{CurtisGPU:2017} have investigated SIMT-based chemical kinetic integration on GPUs, however SIMD-based chemical kinetics have been less studied in comparison.
+\textcite{stone2016} implemented a linearly-implicit fourth-order stiff Rosenbrock solver in the OpenCL for various platforms including the CPU, GPU and the MIC.
+The shallow-SIMD and SIMT-vectorization models were investigated in OpenCL and compared to an OpenMP baseline code that was deep-vectorized by simple compiler hints (a.k.a pragmas).
+The shallow vectorization improved the integrator performance over the OpenMP baseline by \SIrange{2.5}{2.8}{$\times$} on the CPU and \SIrange{4.7}{4.9}{$\times$} on the MIC, while the GPU performance while the GPU was \SIrange{1.4}{1.6}{$\times$} slower than the OpenMP baseline due to thread-divergence concerns.
+\textcite{kroshko2013efficient} implemented a shallow-SIMD vectorized third order stiff Rosenbrock integrator for atmospheric chemistry on a Cell Broadband Engine---a specially designed vector processor---finding a speedup of \SI{1.89}{$\times$} over a serial version of the same code, a \SI{94}{\percent} parallel efficiency.
+
+This work will study the performance of chemical source term evaluation of automatically generated shallow-SIMD vectorized codes for a wide range chemical kinetic models on the CPU.
+The performance will be compared to a baseline SIMT vectorized code to determine the effective SIMD speedup.
 Finally future extensions to this work will be detailed.
@@
 This is a direction that may be explored in the future to compare the speedups of the various compilers.
-
-\subsection{Data Ordering and Vectorization Patterns}
-\label{S:data}
-When storing arrays for a chemical kinetic model, e.g. the concentration of species $k$ for the $j$-th of thermo-chemical state $[C_k]_j$ there are two data-ordering choices.
-In the ``F'' (Fortran column-major) format, the concentrations of a single species $k$ over all $N$ thermo-chemical states are adjacent: $[C_k]_j, [C_k]_{j + 1}\ldots[C_k]_{N}$;
-conversely in the ``C'' (C row-major) format, the species concentrations for the $j$-th thermo-chemical state: $[C_k]_j, [C_{k+1}]_j\ldots$ are stored sequentially in memory;
-This ordering---along with the device (CPU, GPU, etc.) in question---have a large effect on the performance of SIMD\slash SIMT-vectorized algorithms.
-
-In this work a \textit{shallow}-SIMD vectorization, in which each SIMD-lane evaluates chemical source terms for a different thermo-chemical state, has been implemented.
-If the data is stored in ``F''-order, the SIMD-lanes will access sequential locations in memory however the next species' concentration---for example---will be $N$ memory locations away, potentially resulting in more cache-misses on the CPU.
-In future efforts, a so called \textit{deep}-SIMD vectorization will be pursued, in which a vector-unit cooperates to evaluate chemical source terms for a single thermo-chemical state.
-In this context, a deep-vectorization may result in SIMD-\textit{waste}---similar to thread-divergence in GPU-processing---caused by different SIMD-lanes executing different instructions (e.g.
-from differing for-loop bounds).
-However, SIMD-lanes in a ``C''-ordered deep-vectorization access sequential memory locations and further the data locality of e.g., species concentrations, will likely result in improved CPU cache-hit rates.
-Past studies additionally demonstrate promising results for deep-SIMT vectorization on GPUs~\cite{Sewerin20151375}.
-In Sec.~\ref{S:results} the effect of this ordering on the shallow-SIMD vectorization performance will be studied.
 

### 2017-02-17 scripts/general_plotting.py

@@ -1,5 +1,6 @@
 """Module with
 """
+from __future__ import division
 import numpy as np
 import matplotlib.pyplot as plt
 import plot_styles as ps
@@ -30,12 +31,13 @@ def process_data(plotdata, plot, reacs_as_x=True,
     y_vals = []
     err_vals = []
     for run in plotdata:
+        ys = np.array([getattr(data, plot) for data in run.rundata])
+        zs = np.std(ys)
         if plot == 'runtime':
-            run_y = [np.array(getattr(data, plot)) / data.num_conditions for data in run.rundata]
-        else:
-            run_y = [getattr(data, plot) for data in run.rundata]
-        err_vals.append(np.std(run_y))
-        y_vals.append(np.mean(run_y))
+            ys = np.array([getattr(data, plot) / data.num_conditions for data in run.rundata])
+            zs /= data.num_conditions
+        err_vals.append(zs)
+        y_vals.append(np.mean(ys))
     return x_vals, y_vals, err_vals
 
 

### 2017-02-17 scripts/perf_plotter.py

@@ -39,7 +39,8 @@ def __compare(r, name, compare_value, plot_cores=False):
 
 def plotter(plot_name='', show=True, plot_reacs=True, norm=True,
         legend_handler=None, marker_func=None,
-        minx=None, miny=None, maxx=None, maxy=None, **filters):
+        minx=None, miny=None, maxx=None, maxy=None, ylog=False,
+        return_vals=False, **filters):
     data = copy.copy(data_clean)
     #create fig, ax
     fig = plt.figure()
@@ -82,6 +83,7 @@ def plotter(plot_name='', show=True, plot_reacs=True, norm=True,
             plot_reacs = False
             plot_cores = True
 
+    retval = None
     #delete diff for vecwidth / par thing
     if 'vectype' in [diff_check[loc] for loc in diff_locs]:
         ind = next((i for i, loc in enumerate(diff_locs)
@@ -159,6 +161,11 @@ def plotter(plot_name='', show=True, plot_reacs=True, norm=True,
                     y_vals.append(y)
                     z_vals.append(z)
 
+        if return_vals:
+            def __copy_arr(val):
+                return [v[:] for v in val]
+            retval = [__copy_arr(x_vals), __copy_arr(y_vals), __copy_arr(z_vals), copy.copy(labels)]
+
         #second pass - normalize
         if norm and not plot_cores:
             xlen = len(next(x for x in x_vals if x))
@@ -184,6 +191,9 @@ def plotter(plot_name='', show=True, plot_reacs=True, norm=True,
                 x_vals[ix] = x_vals[ix][1:]
                 y_vals[ix] = y_vals[ix][1:]
                 z_vals[ix] = z_vals[ix][1:]
+
+        if ylog:
+            ax.set_yscale('log')
 
         #and finally plot
         for i in range(len(y_vals)):
@@ -212,7 +222,7 @@ def plotter(plot_name='', show=True, plot_reacs=True, norm=True,
         plt.savefig(plot_name)
     if show:
         plt.show()
-    return fig, ax
+    return retval
 
 
 if __name__ == '__main__':

### 2017-02-17 scripts/simd_and_order_comp.py

@@ -41,6 +41,26 @@ handles = [mlines.Line2D([], [], color=SIMT_color, marker=SIMT_sym, linestyle=''
                           label='F-order', markerfacecolor='none', markeredgecolor='k')]
 legend = (handles, labels)
 
-pp.plotter(show=True,
+xval, yval, zval, label = pp.plotter('../SIMD_SIMT_Order_comparison.pdf', show=False,
     cores='1', vecwidth='8', rates='hybrid', kernel='split',
-    legend_handler=legend, marker_func=marker_func)
+    legend_handler=legend, marker_func=marker_func, return_vals=True,
+    norm=False, ylog=True)
+
+#print stats
+C_order = [i for i,l in enumerate(label) if 'C-order' in l]
+F_order = [i for i in range(len(label)) if i not in C_order]
+
+C_SIMT = next(i for i in C_order if 'SIMT' in label[i])
+C_SIMD = next(i for i in C_order if i != C_SIMT)
+F_SIMT = next(i for i in F_order if 'SIMT' in label[i])
+F_SIMD = next(i for i in F_order if i != F_SIMT)
+
+print 'SIMT vs SIMD'
+for i in range(len(yval[0])):
+    print i, 'C', yval[C_SIMT][i] / yval[C_SIMD][i]
+    print i, 'F', yval[F_SIMT][i] / yval[F_SIMD][i]
+
+print 'C vs F'
+for i in range(len(yval[0])):
+    print i, 'SIMT', yval[F_SIMT][i] / yval[C_SIMT][i]
+    print i, 'SIMD', yval[F_SIMD][i] / yval[C_SIMD][i]

### 2017-02-17 paper.tex
minor plotting updates; update results

@@
 In a \textit{deep}-SIMD vectorization, a vector-unit cooperates to evaluate chemical source terms for a single thermo-chemical state, thus SIMD-lanes loading species concentrations $k\dots k+V$ (for a vector width of $V$) will access sequential memory locations if the data is stored in ``C''-order.
-Further, in the ``C''-ordering the furthest difference between any two concentrations in a single thermo-chemical state is at most $N_s \ll N$ (in most cases), greatly improving the data-locality and chance of a cache-hit on the CPU.
-However, a deep-vectorization may result in SIMD-\textit{waste}---similar to thread-divergence in GPU-processing---caused by different SIMD-lanes executing different instructions (e.g.
+Further, in the ``C''-ordering the furthest difference between any two concentrations in a single thermo-chemical state is at most $N_s$, with $N_s \ll N$ in most cases; this greatly improved data-locality increases the chances of a cache-hit on the CPU.
+However, a deep vectorization may result in SIMD-\textit{waste}---similar to thread-divergence in GPU-processing---caused by different SIMD-lanes executing different instructions (e.g.
 from differing for-loop bounds).
@@
 \textcite{stone2016} implemented a linearly-implicit fourth-order stiff Rosenbrock solver in the OpenCL for various platforms including the CPU, GPU and the MIC.
-The shallow-SIMD and SIMT-vectorization models were investigated in OpenCL and compared to an OpenMP baseline code that was deep-vectorized by simple compiler hints (a.k.a pragmas).
+The shallow-SIMD and SIMT-vectorization models were investigated in OpenCL and compared to an OpenMP baseline code that was deep vectorized by simple compiler hints (a.k.a pragmas).
 The shallow vectorization improved the integrator performance over the OpenMP baseline by \SIrange{2.5}{2.8}{$\times$} on the CPU and \SIrange{4.7}{4.9}{$\times$} on the MIC, while the GPU performance while the GPU was \SIrange{1.4}{1.6}{$\times$} slower than the OpenMP baseline due to thread-divergence concerns.
@@
   \centering
-  \begin{subfigure}[t]{0.5\textwidth}
-    \includegraphics[width=\textwidth]{SIMD_SIMT_Order_comparison.pdf}
-    \caption{}
+  \begin{subfigure}[t]{0.48\textwidth}
+    \includegraphics[width=\textwidth]{SIMD_SIMT_Order_comparison}
+    \caption{Runtime in milliseconds per thermo-chemical state of shallow-SIMD and SIMT-vectorization on a single CPU core for both ``C'' (filled symbols) and ``F'' (empty symbols) data ordering.}
     \label{F:SIMDComp}
   \end{subfigure}
-  ~
-
+  \hfill
+  \begin{subfigure}[t]{0.48\textwidth}
+    \includegraphics[width=\textwidth]{SIMD_Vecwidth_comparison}
+    \caption{Speedup of a ``C''-ordered shallow-SIMD vectorization on a single CPU core when varying the supplied vector-width.}
+    \label{F:Veccomp}
+  \end{subfigure} \\
+  \begin{subfigure}[t]{0.5\textwidth}
+    \includegraphics[width=\textwidth]{SIMD_scaling}
+    \caption{Strong parallel scaling of a ``C''-ordered shallow-SIMD vectorization using a vector width of 16.}
+    \label{F:SIMDscale}
+  \end{subfigure}
 \end{figure}
 
-\label{results}
+In Fig.~\ref{F:SIMDComp}, the performance of the shallow-SIMD and SIMT vectorization are compared on a single CPU core for both ``C'' and ``F'' memory formats.
+The ``C''-ordered SIMD\slash SIMT vectorizations show a speedup over the equivalent ``F''-ordered code by \SIrange{1.36}{1.96}{$\times$} and \SIrange{1.60}{2.38}{$\times$} respectively.
+This demonstrates that enhanced data-locality of the ``C''-ordering for CPU caching---discussed in Sec.~\ref{S:data}---is key to high performance on the CPU, and suggests that a deep vectorization would see even greater accelerations.
+The shallow vectorization is consistently faster than the SIMT vectorization, by \SIrange{1.72}{1.88}{$\times$} and \SIrange{2.10}{2.42}{$\times$} for the ``C'' and ``F'' memory formats respectively.
+This is less than the theoretical maximum speedup on this CPU, which can process \SI{256}{\bit} AVX instructions (i.e. four double-precision operations) concurrently.
+Most likely this is due to the non-sequential memory access patterns of the shallow vectorization discussed previously, but could also be caused in part by inefficiencies in translation of the OpenCL code to AVX code, a process handled by the Intel Opencl Runtime.
+
+Figure~\ref{F:Veccomp} investigates the effect of changing the vector-width supplied to a ``C''-ordered shallow-SIMD vectorization on a single CPU core.
+Significant speedups are seen with larger vector-widths for the smaller models, however as the model size increases data must be loaded from memory locations increasingly far apart and the performance of all three vector widths becomes similar.
+Finally in Fig.~\ref{F:SIMDscale} the strong parallel scaling of the chemical source term evaluation is examined for all four models for a ``C''-ordered shallow vectorization with vector-width of 16:
+The scaling efficiency for all models exhibits a power-law dependence, decreasing with the number of cores used.
+Interestingly the efficiency for the hydrogen model decreases fastest, while the larger models have similar efficiencies with isopentanol having the highest efficiency in all cases.
+This suggests that evaluation of the chemical-source terms for a single thermo-chemical state for the smaller models may not saturate the processing power of the CPU, some efficiency may be lost due to context switching within OpenCL itself.
+Additionally, OpenCL allows assigning of work to devices by shared resources e.g., a L3 cache, which was not pursued in this work for simplicity; this should be investigated in future work to determine the effect on scaling efficiency.
 

### 2017-02-17 scripts/perf_plotter.py

@@ -202,7 +202,7 @@ def plotter(plot_name='', show=True, plot_reacs=True, norm=True,
         ax.set_xlim([minx, maxx])
         ax.set_ylim([miny, maxy])
 
-    ylabel = r'Runtime ($\frac{\si{\milli\second}}{\text{state}}$)'
+    ylabel = r'Runtime (\si{\milli\second} / state)'
     xlabel = r'Number of {} in Model'.format('Species' if not plot_reacs else 'Reactions')
     if norm:
         ylabel = 'Speedup'

### 2017-02-17 scripts/plot_styles.py

@@ -13,7 +13,7 @@ plt.rc('text.latex',
 plt.rc('font', family='serif')
 
 legend_style = {'loc':0,
-    'fontsize':16,
+    'fontsize':18,
     'numpoints':1,
     'shadow':True,
     'fancybox':True

### 2017-02-17 scripts/general_plotting.py
fix labels in graph, sort

@@ -7,12 +7,6 @@ import plot_styles as ps
 
 font_size = 'large'
 
-legend_key = {'H2':r'H$_2$/CO',
-              'CH4':r'GRI-Mech 3.0',
-              'C2H4':r'USC-Mech II',
-              'IC5H11OH':r'IC$_5$H$_{11}$OH'
-              }
-
 def process_data(plotdata, plot, reacs_as_x=True,
     plot_cores=False):
     """

### 2017-02-17 scripts/perf_plotter.py

@@ -109,7 +109,12 @@ def plotter(plot_name='', show=True, plot_reacs=True, norm=True,
         try:
             diffs = [sorted(diff, key=lambda x: float(x)) for diff in diffs]
         except:
-            diffs = [sorted(diff) for diff in diffs]
+            if plot_cores:
+                #sort by mech size
+                diffs = [sorted(diff, key=lambda x:
+                    data[x][0].mechdata.num_reactions) for diff in diffs]
+            else:
+                diffs = [sorted(diff) for diff in diffs]
 
         #first pass - process data
         x_vals = []

### 2017-02-17 scripts/plot_styles.py

@@ -46,6 +46,12 @@ class dummy_formatter(object):
         return next(self.values[i] for i, x in enumerate(self.choices)
             if x == val)
 
+legend_key = {'H2':r'H$_2$/CO',
+              'CH4':r'GRI-Mech 3.0',
+              'C2H4':r'USC-Mech II',
+              'IC5H11OH':r'IC$_5$H$_{11}$OH'
+              }
+
 def pretty_names(pname, short=False):
     pname_dict = {'runtime' : 'Runtime',
         'comptime' : 'Compilation time',
@@ -61,7 +67,7 @@ def pretty_names(pname, short=False):
         'rates' : dummy_formatter({'fixed' : 'Fixed Rate Specialization',
             'hybrid' : 'Hybrid Rate Specialization',
             'full' : 'Full Rate Specialization'}),
-        'mechdata.mech' : r'\ce{{{}}}' #used in core plotting
+        'mechdata.mech' : dummy_formatter(legend_key)
     }
     if pname in pname_dict:
         return pname_dict[pname]

### 2017-02-17 paper.tex
fix labels in graph, sort
add first conclusions

@@
 data ordering, vectorization, threading patterns etc.
-Additionally, \texttt{loo.py} allows for code generation for multiple languages, including (non-vectorized) C, CUDA, and Intel's ISPC compiler, which targets SIMD vectorization on the CPU and MIC.
-This is a direction that may be explored in the future to compare the speedups of the various compilers.
 
@@
 \toprule
-\multicolumn{1}{c}{Model} & \multicolumn{1}{c}{$E_{R_{i}^{\prime}}$} & \multicolumn{1}{c}{$E_{R_{i}^{\prime\prime}}$} &\multicolumn{1}{c}{$E_{R_{i}}$} & \multicolumn{1}{c}{$\text{P}_{\text{min}}^{*}\si{\percent}$} & \multicolumn{1}{c}{$E_{\frac{\text{d}[C_k]}{\text{d}t}}$} & \multicolumn{1}{c}{$E_{\frac{\text{d}[T]}{\text{d}t}}$} \\
+\multicolumn{1}{l}{Model} & \multicolumn{1}{c}{$E_{R_{i}^{\prime}}$} & \multicolumn{1}{c}{$E_{R_{i}^{\prime\prime}}$} &\multicolumn{1}{c}{$E_{R_{i}}$} & \multicolumn{1}{c}{$\text{P}_{\text{min}}^{*}\si{\percent}$} & \multicolumn{1}{c}{$E_{\frac{\text{d}[C_k]}{\text{d}t}}$} & \multicolumn{1}{c}{$E_{\frac{\text{d}[T]}{\text{d}t}}$} \\
 \midrule
-\multicolumn{1}{c}{\ce{H2}\slash~\ce{CO}} & 1.20e-8 & 6.33e-8 & 1.32e1 & 4.66e6 & 2.37e1 & 2.12e5 \\
-\multicolumn{1}{c}{GRI-Mech.~3.0}  & 3.07e-8 & 5.90e-8 & 1.21e0 & 2.67e04 & 2.84e0 & 2.64e4 \\
-\multicolumn{1}{c}{USC-Mech II}  & 1.07e-7 & 1.37e-7 & 2.39e0 & 4.12e02 & 7.92e00 & 3.22e03 \\
+\multicolumn{1}{l}{\ce{H2}\slash~\ce{CO}\cite{Burke:2011fh}} & 1.20e-8 & 6.33e-8 & 1.32e1 & 4.66e6 & 2.37e1 & 2.12e5 \\
+\multicolumn{1}{l}{GRI-Mech.~3.0\cite{smith_gri-mech_30}}  & 3.07e-8 & 5.90e-8 & 1.21e0 & 2.67e04 & 2.84e0 & 2.64e4 \\
+\multicolumn{1}{l}{USC-Mech II\cite{Wang:2007}}  & 1.07e-7 & 1.37e-7 & 2.39e0 & 4.12e02 & 7.92e00 & 3.22e03 \\
+\multicolumn{1}{l}{\ce{IC5H11OH}\cite{Sarathy:2013jr}} & -1 &  -1 & -1 & -1 & -1 & -1 \\
 \bottomrule
@@
 \section{Conclusions}
-%
+In this work, a shallow SIMD-vectorized chemical source term evaluation OpenCL code generator was developed and validated against Cantera~\cite{Cantera} for a wide range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr}.
+Significant speedups of up to \SI{2.42}{$\times$} over a baseline SIMT-vectorized code were observed.
+Two data-ordering schemes were investigated, showing a clear performance benefit for use of the ``C'' (row-major) ordering for the shallow vectorized code.
+Further, this study suggests that a deep SIMD-vectorized code---currently under development---may see even greater accelerations due to expected increases in cache-hit rates.
+Larger vector widths were found to provide accelerations for the smaller models studied, but performance was similar over all vector widths tested for the largest models tested.
+Finally, the strong parallel scaling efficiency was examined for the shallow vectorized code; the larger models exhibited higher scaling efficiency, possibly due to failure of the \ce{H2}\slash\ce{CO} model to saturate the CPU throughput.
+Future extensions of this work will include development of a SIMD\slash SIMT-accelerated sparse-analytical Jacobian code targeted at CPUs, GPUs and MICs.
 
@@
 
-\noindent\textbf{Page Limits:} The total length of the paper including references should be limited to 6 pages.
-
 \printbibliography[heading=bibintoc]

### 2017-02-17 paper.tex
updates

@@
 
-\author[1]{Author Name}
-\author[1]{Author Name}
-\author[2]{Author Name}
-\author[2,*]{Author Name}
+\author[1,*]{Nicholas Curtis}
+\author[1]{Chih-Jen Sung}
 
-\affil[1]{Department, Institution, Address, Country}
-\affil[2]{Department, Institution, Address, Country}
-\affil[*]{Corresponding author: \email{author@university.edu}}
+\affil[1]{Department of Mechanical Engineering, University of Connecticut, Storrs, CT 06269, USA}
+\affil[*]{Corresponding author: \email{nicholas.curtis@uconn.edu}}
 
@@
 \begin{abstract} % not to exceed 200 words
-Abstract should be between 150--200 words and should state briefly the purpose
-of the research, the principal results and major conclusions. An abstract is
-often presented separately from the article, so it must be able to stand alone.
-For this reason, References should be avoided, but if essential, then cite the
-author(s) and year(s). Also, non-standard or uncommon abbreviations should be
-avoided, but if essential they must be defined at their first mention in the
-abstract itself.
+A code generation platform for single-instruction, multiple-data vectorized OpenCL chemical source term evaluation was developed and validated against Cantera for a wide range of chemical kinetic models.
+Speedups of \SI{1.88}{2.42}{$\times$} were observed over a baseline single-instruction, multiple-thread parallelized code showing a clear performance benefit to use of a row-major data layout.
+Larger vector widths were found to produce speedups up to \SI{1.8}{$\times$} for smaller models, but had similar performance to smaller vector widths as the model size increased.
+Finally, the parallel scaling efficiency was found to be slightly better for larger models, and future extensions of this work were discussed.
 \end{abstract}
@@
 \begin{keyword}
-    Keyword1\sep Keyword2\sep Keyword3\sep Keyword4
+    Chemical Kinetics\sep Single Instruction, Multiple Data\sep Vector Processing
 \end{keyword}
@@
   \centering
-  \begin{subfigure}[t]{0.54\linewidth}
+  \begin{subfigure}[t]{0.45\linewidth}
       \includegraphics[width=\textwidth]{SIMD.pdf}
@@
   \hfill
-  \begin{subfigure}[t]{0.43\linewidth}
+  \begin{subfigure}[t]{0.45\linewidth}
       \includegraphics[width=\textwidth]{mem_access.pdf}
@@
 Finally in Fig.~\ref{F:SIMDscale} the strong parallel scaling of the chemical source term evaluation is examined for all four models for a ``C''-ordered shallow vectorization with vector-width of 16:
-The scaling efficiency for all models exhibits a power-law dependence, decreasing with the number of cores used.
-Interestingly the efficiency for the hydrogen model decreases fastest, while the larger models have similar efficiencies with isopentanol having the highest efficiency in all cases.
+The scaling efficiency for decreases with the number of cores used for all models; interestingly the efficiency for the hydrogen model decreases fastest, while the larger models have similar efficiencies with isopentanol having the highest efficiency in all cases.
 This suggests that evaluation of the chemical-source terms for a single thermo-chemical state for the smaller models may not saturate the processing power of the CPU, some efficiency may be lost due to context switching within OpenCL itself.
@@
 In this work, a shallow SIMD-vectorized chemical source term evaluation OpenCL code generator was developed and validated against Cantera~\cite{Cantera} for a wide range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr}.
-Significant speedups of up to \SI{2.42}{$\times$} over a baseline SIMT-vectorized code were observed.
+Significant speedups of up to \SIrange{1.88}{2.42}{$\times$} over a baseline SIMT-vectorized code were observed.
 Two data-ordering schemes were investigated, showing a clear performance benefit for use of the ``C'' (row-major) ordering for the shallow vectorized code.

### 2017-02-17 paper.tex
pass 1

@@
 \begin{abstract} % not to exceed 200 words
-A code generation platform for single-instruction, multiple-data vectorized OpenCL chemical source term evaluation was developed and validated against Cantera for a wide range of chemical kinetic models.
-Speedups of \SI{1.88}{2.42}{$\times$} were observed over a baseline single-instruction, multiple-thread parallelized code showing a clear performance benefit to use of a row-major data layout.
+A code generation platform for single-instruction, multiple-data-vectorized (SIMD) chemical source term evaluation was developed and validated against Cantera for a wide range of chemical kinetic models.
+Speedups of up to \SIrange{1.88}{2.42}{$\times$} were observed over a baseline single-instruction, multiple-thread (SIMT) parallelized code.
+Additionally, use of a row-major data layout was found to increase both the SIMD and SIMT performance by up to \SIrange{1.96}{2.38}{$\times$} over a column-major data layout.
 Larger vector widths were found to produce speedups up to \SI{1.8}{$\times$} for smaller models, but had similar performance to smaller vector widths as the model size increased.
@@
 \begin{keyword}
-    Chemical Kinetics\sep Single Instruction, Multiple Data\sep Vector Processing
+    Chemical Kinetics\sep SIMD \sep Vector Processing
 \end{keyword}
@@
 Single Instruction, Multiple-Data (SIMD) processing, also known as vector-processing, is an important parallel computing paradigm used increasingly in scientific computing.
-While central processing unit (CPU) clock speeds have increased rapidly over the past few decades---the phenomena known as Moore's Law---power consumption and head dissipation issues have caused this trend to slow.
+While central processing unit (CPU) clock speeds have increased rapidly over the past few decades---the phenomena known as Moore's Law---power consumption and heat dissipation issues have caused this trend to slow.
 Traditional multi-core parallelism is often used to increase CPU performance, however SIMD processors---and the related Single Instruction, Multiple Thread (SIMT) processors, e.g.
@@
 A SIMD instruction utilizes specialized vector processing hardware to execute the same floating point operation (e.g.
-multiplication, division, etc.) on multiple pieces of data concurrently (Fig.~\ref{F:SIMD}).
-Nearly all modern CPUs have vector processing units capable of running SIMD instructions, typically capable of handling 2--4 concurrent double precision operations---known as the vector width.
+multiplication, division, etc.) on a several different data concurrently (Fig.~\ref{F:SIMD}); the number of concurrent operations is known as the vector-width.
+Nearly all modern CPUs have vector processing units capable of running SIMD instructions, typically capable of handling 2--4 concurrent double precision operations.
 Additionally, specialized hardware accelerators, e.g.
@@
 In Fig.~\ref{F:mem} this would correspond to storing $[C_0]_0$ in index \num{0}, $[C_0]_1$ in index \num{1} and so on, with $N_1 = N$, and $N_2 = N_s$, the number of species in the model.
-In the ``C'' (C row-major) format, the storage is flipped with $[C_0]_0$ in index \num{0}, $[C_1]_0$ in index \num{1} etc., with $N_1 = N_s$ and $N_2 = N$.
+In the ``C'' (C row-major) format, the storage order is transposed with $[C_0]_0$ in index \num{0}, $[C_1]_0$ in index \num{1} etc., and $N_1 = N_s$, $N_2 = N$.
 This ordering---along with the device (CPU, GPU, etc.) in question---have a large effect on the performance of SIMD\slash SIMT-vectorized algorithms.
 
-In a \textit{shallow}-SIMD vectorization, each SIMD-lane evaluates chemical source terms for a different thermo-chemical state, has been implemented.
-If the data is stored in ``F''-order, the SIMD-lanes accessing the concentration of species $k$ for different states will load sequential memory locations in Fig.~\ref{F:mem} in memory however the $k+1$'th species concentration will be $N$ memory locations away, increasing the likelihood of cache-misses on the CPU.
-In a \textit{deep}-SIMD vectorization, a vector-unit cooperates to evaluate chemical source terms for a single thermo-chemical state, thus SIMD-lanes loading species concentrations $k\dots k+V$ (for a vector width of $V$) will access sequential memory locations if the data is stored in ``C''-order.
+In a \textit{shallow}-SIMD vectorization, each SIMD-lane evaluates chemical source terms for a different thermo-chemical state.
+If the data is stored in ``F''-order, the SIMD-lanes accessing the concentration of species $k$ for states $j, j+1\ldots$ will load sequential locations in memory; the $k+1$th species concentration will be $N$ memory locations away however, increasing the likelihood of cache-misses on the CPU.
+In a \textit{deep}-SIMD vectorization, a vector-unit cooperates to evaluate chemical source terms for a single thermo-chemical state $j$, thus SIMD-lanes loading species concentrations $k, k+1\ldots$ will access sequential memory locations if the data is stored in ``C''-order.
 Further, in the ``C''-ordering the furthest difference between any two concentrations in a single thermo-chemical state is at most $N_s$, with $N_s \ll N$ in most cases; this greatly improved data-locality increases the chances of a cache-hit on the CPU.
@@
 from differing for-loop bounds).
-In Sec.~\ref{S:results} the effect of this ordering on the shallow-SIMD vectorization performance will be studied.
 
-\subsection{Previous works and goals of this tudy}
+Finally, in a SIMT-vectorization, threads execute the same instruction over multiple data in lock-step without use of specialized vectorization instructions.
+This is the paradigm used in GPU-processing, a more through review of which can be found in our previous work~\cite{CurtisGPU}; one key-exception in this case is that the algorithm executed to evaluate the chemical source terms is identical for all thermo-chemical states, thus there is no thread-divergence for a shallow SIMT-vectorization.
+In Sec.~\ref{S:results} the effect of the data-ordering on the shallow-SIMD\slash SIMT vectorization performance will be studied.
+
+\subsection{Previous works and goals of this study}
 Several recent works, e.g. \cite{CurtisGPU:2017} have investigated SIMT-based chemical kinetic integration on GPUs, however SIMD-based chemical kinetics have been less studied in comparison.
 \textcite{stone2016} implemented a linearly-implicit fourth-order stiff Rosenbrock solver in the OpenCL for various platforms including the CPU, GPU and the MIC.
-The shallow-SIMD and SIMT-vectorization models were investigated in OpenCL and compared to an OpenMP baseline code that was deep vectorized by simple compiler hints (a.k.a pragmas).
-The shallow vectorization improved the integrator performance over the OpenMP baseline by \SIrange{2.5}{2.8}{$\times$} on the CPU and \SIrange{4.7}{4.9}{$\times$} on the MIC, while the GPU performance while the GPU was \SIrange{1.4}{1.6}{$\times$} slower than the OpenMP baseline due to thread-divergence concerns.
+The shallow-SIMD and SIMT-vectorization models were implemented in OpenCL and compared to an OpenMP baseline code that was deep SIMD-vectorized by simple compiler hints (a.k.a pragmas).
+The shallow-SIMD vectorization improved the integrator performance over the OpenMP baseline by \SIrange{2.5}{2.8}{$\times$} on the CPU and \SIrange{4.7}{4.9}{$\times$} on the MIC, while the GPU performance was \SIrange{1.4}{1.6}{$\times$} slower than the OpenMP baseline due to thread-divergence concerns.
 \textcite{kroshko2013efficient} implemented a shallow-SIMD vectorized third order stiff Rosenbrock integrator for atmospheric chemistry on a Cell Broadband Engine---a specially designed vector processor---finding a speedup of \SI{1.89}{$\times$} over a serial version of the same code, a \SI{94}{\percent} parallel efficiency.
@@
 This work will study the performance of chemical source term evaluation of automatically generated shallow-SIMD vectorized codes for a wide range chemical kinetic models on the CPU.
-The performance will be compared to a baseline SIMT vectorized code to determine the effective SIMD speedup.
+The performance will be compared to a baseline shallow-SIMT vectorized code to determine the effective SIMD speedup.
 Finally future extensions to this work will be detailed.
@@
 
-In addition, the temperature rate of change using the constant-pressure assumption\footnote{Note: in this context, the ``constant-pressure assumption'' refers to evaluation within a reaction sub-step in the operator splitting scheme, rather than a general constant-pressure reactive-flow simulation.} is evaluated as:
+In addition, the temperature rate of change using the constant-pressure assumption\footnote{Note: in this context, the ``constant-pressure assumption'' refers to evaluation within a reaction sub-step in the operator splitting scheme, rather than a general constant-pressure reactive-flow simulation.} is evaluated on a molar basis as:
 \begin{equation}
@@
 \end{equation}
-although the code is equally capable of a evaluating the temperature rate using a constant-volume assumption, we omit this here for brevity.
+where $H^{\degree}_k$ and $c_{p_k}^{\degree}$ are the molar specific enthalpy and specific heat respectively.
+Although the code is equally capable of a evaluating the temperature rate using a constant-volume assumption, we omit this here for brevity.
 
 \subsection{Code Generation}
-Code generation is handled by the python package \texttt{loo.py}~\cite{kloeckner_loopy_2014}, which provides an interface to OpenCL allowing for unit testing and changes of program structure, e.g.
+Code generation is enabled by the python package \texttt{loo.py}~\cite{kloeckner_loopy_2014}, which translates user specified psuedo-code and data to OpenCL, allowing for unit testing and easy changes of program structure, e.g.
 data ordering, vectorization, threading patterns etc.
@@
 
-In computing the net rate of progress of reaction $i$ from the forward and reverse ROP: $R_{i} = R_{i}^{\prime} - R_{i}^{\prime\prime}$, precision can easily be lost as the net rate of progress may be---particularly near chemical equilibrium---many orders of magnitude smaller than the forward or reverse rates.
+In computing the net ROP of reaction $i$ from the forward and reverse ROP: $R_{i} = R_{i}^{\prime} - R_{i}^{\prime\prime}$, precision can easily be lost as the net ROP may be---particularly near chemical equilibrium---many orders of magnitude smaller than the forward or reverse rates.
 A base-ten bound on loss of precision~\cite{goldberg1991every} can be found as:
 \begin{equation}
-\text{P}_{\text{min},i} = 2^{-p_i} \le \left\lvert 1 - \frac{R_{i}^{\prime}}{R_{i}^{\prime\prime}} \right\rvert \le 2^{-q_i} = \text{P}_{\text{max},i}
+\text{L}_{i,\text{min}} = 2^{-p_i} \le \left\lvert 1 - \frac{R_{i}^{\prime}}{R_{i}^{\prime\prime}} \right\rvert \le 2^{-q_i} = \text{L}_{i,\text{max}}}
 \label{e:bounds}
@@
 with $q_i \le p_i$.
-To obtain an estimate of the a-priori precision bounds ($\text{P}_{\text{min},i}$ and $\text{P}_{\text{max},i}$), the forward and reverse ROP from Cantera are substituted into Eq.~\eqref{e:bounds}.
+To obtain an estimate of the precision bounds, the forward and reverse ROP from Cantera are substituted into Eq.~\eqref{e:bounds}.
 Additionally a higher-precision number format (in this case, the Python Decimal class) is used in this computation to maintain accuracy.
-Finally, when determining the error in the net ROP between this code and Cantera, the bounds on the percent error attributable to precision loss, $\text{P}_{\text{err},i}$ can be estimated as:
+Finally, when determining the error in the net ROP between this code and Cantera, the bounds on the percent error attributable to precision loss, $\text{P}_{i,\text{err}}$ can be estimated as:
 \begin{equation}
-\text{P}_{i,\text{min}} = 100 * \frac{\text{P}_{\text{min},i}}{\left\lvert\ R_{i,\text{CT}} - R_{i}\right\rvert} \le
-\text{P}_{i} \le 100 * \frac{\text{P}_{\text{max},i}}{\left\lvert R_{i,\text{CT}} - R_{i} \right\rvert} = \text{P}_{i,\text{max}}
+\text{P}_{i,\text{min}} = 100 * \frac{\text{L}_{i,\text{min}}}{\left\lvert\ R_{i,\text{CT}} - R_{i}\right\rvert} \le
+\text{P}_{i,\text{err}} \le 100 * \frac{\text{L}_{i,\text{max}}}{\left\lvert R_{i,\text{CT}} - R_{i} \right\rvert} = \text{P}_{i,\text{max}}
 \label{e:rel_bound}
@@
 \end{equation}
-and the lower precision bound for the net rate of progress error norm is set to $P_{\text{min}}^{*} = P_{i,\text{min}}$ for the $i, j$ value that results in the maximum error $E_{R_i}$.
+and the lower precision bound for the net ROP error norm is set to $\text{P}_{\text{min}}^{*} = \text{P}_{i,\text{min}}$ for the $i, j$ value that results in the maximum error $E_{R_i}$.
 
@@
 In Table~\ref{T:error}, we see the results of this code as compared to Cantera on a library of partially stirred reaction conditions (PaSR) described in our previous works~\cite{CurtisGPU:2017,Niemeyer:2016aa}.
-Very close agreement is observed for the forward and reverse ROP for all models, however the net rate of progress is \numrange{7}{8} orders of magnitude larger.
+Very close agreement is observed for the forward and reverse ROP for all models, however the net ROP is \numrange{7}{8} orders of magnitude larger.
 We see here that the minimum estimated loss of precision is \SIrange{4}{40}{$\times$} the error in net ROP, explaining this jump.
 The species rate error norm is similar in magnitude to that of the net ROP, however the temperature rate error again increases as it depends directly on the species rates.
-We note that the above discussion does not imply that computation of the net ROP will cause large error in chemical kinetic integration---either in this code or Cantera---as this loss of precision only occurs when the net ROP are many orders of magnitude smaller than the forward or reverse rates, implying the reaction is in near-equilibrium.
+We note that the above discussion does not imply that computation of the net ROP will cause large error in chemical kinetic integration---either in this code or Cantera---as this loss of precision only occurs when the forward and reverse ROP are nearly equal, implying the reaction is in near-equilibrium.
 
@@
 
-Figure~\ref{F:Veccomp} investigates the effect of changing the vector-width supplied to a ``C''-ordered shallow-SIMD vectorization on a single CPU core.
-Significant speedups are seen with larger vector-widths for the smaller models, however as the model size increases data must be loaded from memory locations increasingly far apart and the performance of all three vector widths becomes similar.
-Finally in Fig.~\ref{F:SIMDscale} the strong parallel scaling of the chemical source term evaluation is examined for all four models for a ``C''-ordered shallow vectorization with vector-width of 16:
-The scaling efficiency for decreases with the number of cores used for all models; interestingly the efficiency for the hydrogen model decreases fastest, while the larger models have similar efficiencies with isopentanol having the highest efficiency in all cases.
+The effect of changing the vector-width supplied to a ``C''-ordered shallow-SIMD vectorization on a single CPU core is investigated in Fig.~\ref{F:Veccomp}.
+Significant speedups are seen with larger vector-widths for the smaller models, however as the model size increases data must be loaded from increasingly distant memory locations and the performance of all three vector widths becomes similar.
+Finally in Fig.~\ref{F:SIMDscale} the strong parallel scaling of the chemical source term evaluation is examined for all four models with a ``C''-ordered shallow SIMD-vectorization and vector-width of 16:
+Interestingly the efficiency for the hydrogen model decreases fastest, while the larger models have similar efficiencies; the isopentanol model has the highest efficiency in all cases.
 This suggests that evaluation of the chemical-source terms for a single thermo-chemical state for the smaller models may not saturate the processing power of the CPU, some efficiency may be lost due to context switching within OpenCL itself.
-Additionally, OpenCL allows assigning of work to devices by shared resources e.g., a L3 cache, which was not pursued in this work for simplicity; this should be investigated in future work to determine the effect on scaling efficiency.
+Additionally, OpenCL allows assigning of work to devices by shared memory resources e.g., a L3 cache, which was not pursued in this study for simplicity; this should be investigated in future work to determine the effect on scaling efficiency.
 

### 2017-02-17 paper.tex
pass 2

@@
 
-\title{ Title of Paper }
+\title{SIMD-vectorized Chemical Source Term Evaluation}
 
@@
 Single Instruction, Multiple-Data (SIMD) processing, also known as vector-processing, is an important parallel computing paradigm used increasingly in scientific computing.
-While central processing unit (CPU) clock speeds have increased rapidly over the past few decades---the phenomena known as Moore's Law---power consumption and heat dissipation issues have caused this trend to slow.
 Traditional multi-core parallelism is often used to increase CPU performance, however SIMD processors---and the related Single Instruction, Multiple Thread (SIMT) processors, e.g.
@@
 A SIMD instruction utilizes specialized vector processing hardware to execute the same floating point operation (e.g.
-multiplication, division, etc.) on a several different data concurrently (Fig.~\ref{F:SIMD}); the number of concurrent operations is known as the vector-width.
+multiplication, division, etc.) on a several different data concurrently (Fig.~\ref{F:SIMD}); the number of concurrent operations is known as the vector-width\footnote{OpenCL allows for use of vector-widths not equal to the hardware vector-width via implicit conversion, and may provide some performance benefit as will be studied here}.
 Nearly all modern CPUs have vector processing units capable of running SIMD instructions, typically capable of handling 2--4 concurrent double precision operations.
@@
 In a \textit{shallow}-SIMD vectorization, each SIMD-lane evaluates chemical source terms for a different thermo-chemical state.
-If the data is stored in ``F''-order, the SIMD-lanes accessing the concentration of species $k$ for states $j, j+1\ldots$ will load sequential locations in memory; the $k+1$th species concentration will be $N$ memory locations away however, increasing the likelihood of cache-misses on the CPU.
+If the data is stored in ``F''-order, the SIMD-lanes accessing the concentration of species $k$ for states $j, j+1\ldots$ will load sequential locations in memory, however the $k+1$th species concentration will be $N$ memory locations away increasing the likelihood of cache-misses on the CPU.
 In a \textit{deep}-SIMD vectorization, a vector-unit cooperates to evaluate chemical source terms for a single thermo-chemical state $j$, thus SIMD-lanes loading species concentrations $k, k+1\ldots$ will access sequential memory locations if the data is stored in ``C''-order.
@@
 Finally, in a SIMT-vectorization, threads execute the same instruction over multiple data in lock-step without use of specialized vectorization instructions.
-This is the paradigm used in GPU-processing, a more through review of which can be found in our previous work~\cite{CurtisGPU}; one key-exception in this case is that the algorithm executed to evaluate the chemical source terms is identical for all thermo-chemical states, thus there is no thread-divergence for a shallow SIMT-vectorization.
+This is the paradigm used in GPU-processing, a more through review of which can be found in our previous work~\cite{CurtisGPU:2017}; one key-exception in this case is that the algorithm executed to evaluate the chemical source terms is identical for all thermo-chemical states, thus there is no thread-divergence for a shallow SIMT-vectorization.
 In Sec.~\ref{S:results} the effect of the data-ordering on the shallow-SIMD\slash SIMT vectorization performance will be studied.
@@
 
-In addition, the temperature rate of change using the constant-pressure assumption\footnote{Note: in this context, the ``constant-pressure assumption'' refers to evaluation within a reaction sub-step in the operator splitting scheme, rather than a general constant-pressure reactive-flow simulation.} is evaluated on a molar basis as:
+In addition, the temperature rate of change using the constant-pressure assumption\footnote{Note: in this context, the ``constant-pressure assumption'' refers to evaluation within a reaction sub-step in the operator splitting scheme, rather than a general constant-pressure reactive-flow simulation.} is evaluated as:
 \begin{equation}
@@
 In Table~\ref{T:error}, we see the results of this code as compared to Cantera on a library of partially stirred reaction conditions (PaSR) described in our previous works~\cite{CurtisGPU:2017,Niemeyer:2016aa}.
-Very close agreement is observed for the forward and reverse ROP for all models, however the net ROP is \numrange{7}{8} orders of magnitude larger.
+Very close agreement is observed for the forward and reverse ROP for all models, however the net ROP error norm is \numrange{7}{8} orders of magnitude larger.
 We see here that the minimum estimated loss of precision is \SIrange{4}{40}{$\times$} the error in net ROP, explaining this jump.

### 2017-02-21 paper.tex
remove #'s in affil
reorder

@@
 
-\author[1,*]{Nicholas Curtis}
-\author[1]{Chih-Jen Sung}
+\author[*]{Nicholas Curtis}
+\author[ ]{Chih-Jen Sung}
 
-\affil[1]{Department of Mechanical Engineering, University of Connecticut, Storrs, CT 06269, USA}
+\affil[ ]{Department of Mechanical Engineering, University of Connecticut, Storrs, CT 06269, USA}
 \affil[*]{Corresponding author: \email{nicholas.curtis@uconn.edu}}
@@
 \begin{abstract} % not to exceed 200 words
-A code generation platform for single-instruction, multiple-data-vectorized (SIMD) chemical source term evaluation was developed and validated against Cantera for a wide range of chemical kinetic models.
+A code generation platform for single-instruction, multiple-data (SIMD) vectorized chemical source term evaluation was developed and validated against Cantera for a wide range of chemical kinetic models.
 Speedups of up to \SIrange{1.88}{2.42}{$\times$} were observed over a baseline single-instruction, multiple-thread (SIMT) parallelized code.

### 2017-02-21 paper.tex
point out speedup compared to what

@@
 Additionally, use of a row-major data layout was found to increase both the SIMD and SIMT performance by up to \SIrange{1.96}{2.38}{$\times$} over a column-major data layout.
-Larger vector widths were found to produce speedups up to \SI{1.8}{$\times$} for smaller models, but had similar performance to smaller vector widths as the model size increased.
+Larger vector-widths were found to produce speedups up to \SI{1.8}{$\times$} as compared to a small vector-width for smaller models, e.g.
+GRI-Mech 3.0, but as the model size increased the performance of all vector-widths were similar.
 Finally, the parallel scaling efficiency was found to be slightly better for larger models, and future extensions of this work were discussed.

### 2017-02-21 paper.tex
hyphens, cpu defn

@@
 
-Single Instruction, Multiple-Data (SIMD) processing, also known as vector-processing, is an important parallel computing paradigm used increasingly in scientific computing.
-Traditional multi-core parallelism is often used to increase CPU performance, however SIMD processors---and the related Single Instruction, Multiple Thread (SIMT) processors, e.g.
+Single-Instruction, Multiple-Data (SIMD) processing, also known as vector-processing, is an important parallel computing paradigm used increasingly in scientific computing.
+Traditional multi-core parallelism is often used to increase central processing unit (CPU) performance, however SIMD processors---and the related Single-Instruction, Multiple-Thread (SIMT) processors, e.g.
 graphics processing units (GPUs)---have gained recognition due to their increased floating operation throughput.

### 2017-02-21 paper.tex
vecwidth hyphens

@@
 Additionally, specialized hardware accelerators, e.g.
-Intel's Xeon Phi co-processor (or MIC), have been developed that have tens of cores with very wide vector widths (e.g.
-4--8 double precision operations); these very wide vector widths are also expected on forthcoming Intel CPUs (the Skylake Xeon and Cannon Lake architectures).
+Intel's Xeon Phi co-processor (or MIC), have been developed that have tens of cores with very wide vector-widths (e.g.
+4--8 double precision operations); these very wide vector-widths are also expected on forthcoming Intel CPUs (the Skylake Xeon and Cannon Lake architectures).
 
@@
     \includegraphics[width=\textwidth]{SIMD_scaling}
-    \caption{Strong parallel scaling of a ``C''-ordered shallow-SIMD vectorization using a vector width of 16.}
+    \caption{Strong parallel scaling of a ``C''-ordered shallow-SIMD vectorization using a vector-width of 16.}
     \label{F:SIMDscale}
@@
 The effect of changing the vector-width supplied to a ``C''-ordered shallow-SIMD vectorization on a single CPU core is investigated in Fig.~\ref{F:Veccomp}.
-Significant speedups are seen with larger vector-widths for the smaller models, however as the model size increases data must be loaded from increasingly distant memory locations and the performance of all three vector widths becomes similar.
+Significant speedups are seen with larger vector-widths for the smaller models, however as the model size increases data must be loaded from increasingly distant memory locations and the performance of all three vector-widths becomes similar.
 Finally in Fig.~\ref{F:SIMDscale} the strong parallel scaling of the chemical source term evaluation is examined for all four models with a ``C''-ordered shallow SIMD-vectorization and vector-width of 16:
@@
 Further, this study suggests that a deep SIMD-vectorized code---currently under development---may see even greater accelerations due to expected increases in cache-hit rates.
-Larger vector widths were found to provide accelerations for the smaller models studied, but performance was similar over all vector widths tested for the largest models tested.
+Larger vector-widths were found to provide accelerations for the smaller models studied, but performance was similar over all vector-widths tested for the largest models tested.
 Finally, the strong parallel scaling efficiency was examined for the shallow vectorized code; the larger models exhibited higher scaling efficiency, possibly due to failure of the \ce{H2}\slash\ce{CO} model to saturate the CPU throughput.

### 2017-02-21 paper.tex
move to separate figures, update footnote and mic defn update fig

@@
   \centering
-  \begin{subfigure}[t]{0.45\linewidth}
+  \begin{minipage}[t]{0.45\linewidth}
       \includegraphics[width=\textwidth]{SIMD.pdf}
@@
       \label{F:SIMD}
-  \end{subfigure}
+  \end{minipage}
   \hfill
-  \begin{subfigure}[t]{0.45\linewidth}
+  \begin{minipage}[t]{0.45\linewidth}
       \includegraphics[width=\textwidth]{mem_access.pdf}
@@
       \label{F:mem}
-  \end{subfigure}
+  \end{minipage}
 \end{figure}
@@
 A SIMD instruction utilizes specialized vector processing hardware to execute the same floating point operation (e.g.
-multiplication, division, etc.) on a several different data concurrently (Fig.~\ref{F:SIMD}); the number of concurrent operations is known as the vector-width\footnote{OpenCL allows for use of vector-widths not equal to the hardware vector-width via implicit conversion, and may provide some performance benefit as will be studied here}.
+multiplication, division, etc.) on a several different data concurrently (Fig.~\ref{F:SIMD}); the number of concurrent operations is known as the vector-width\footnote{OpenCL allows for use of vector-widths different from the actual hardware vector-width via implicit conversion, and may provide some performance benefit as will be studied in Sec.~\ref{S:results}}.
 Nearly all modern CPUs have vector processing units capable of running SIMD instructions, typically capable of handling 2--4 concurrent double precision operations.
 Additionally, specialized hardware accelerators, e.g.
-Intel's Xeon Phi co-processor (or MIC), have been developed that have tens of cores with very wide vector-widths (e.g.
+Intel's Xeon Phi co-processor or multi integrated core architecture (MIC), have been developed that have tens of cores with very wide vector-widths (e.g.
 4--8 double precision operations); these very wide vector-widths are also expected on forthcoming Intel CPUs (the Skylake Xeon and Cannon Lake architectures).

### 2017-02-23 paper.tex
first pass at dr. sungs edits

@@
 A single processing element (e.g.
-a CPU core) contains a vector unit with several lanes (L0, L1, etc.).
+a CPU core) contains a vector unit with $N$ lanes ($\text{L}_0,\ldots \text{L}_N$).
 The vector unit executes a single instruction concurrently on multiple data.}
@@
       \includegraphics[width=\textwidth]{mem_access.pdf}
-      \caption{Representation of an 2-D array with dimension $N_1 \times N_2$ in CPU memory, with linear index marked.}
+      \caption{Representation of an 2-D array with dimension $N \times K$ in CPU memory, with linear index marked.}
       \label{F:mem}
@@
 \label{S:data}
-When storing arrays for a chemical kinetic model, e.g. the concentration of species $k$ for the $j$-th of thermo-chemical state $[C_k]_j$ there are two data-ordering choices.
-In the ``F'' (Fortran column-major) format, the concentrations of a single species $k$ over all $N$ thermo-chemical states are adjacent.
-In Fig.~\ref{F:mem} this would correspond to storing $[C_0]_0$ in index \num{0}, $[C_0]_1$ in index \num{1} and so on, with $N_1 = N$, and $N_2 = N_s$, the number of species in the model.
-In the ``C'' (C row-major) format, the storage order is transposed with $[C_0]_0$ in index \num{0}, $[C_1]_0$ in index \num{1} etc., and $N_1 = N_s$, $N_2 = N$.
+When storing arrays for a chemical kinetic model, e.g.
+the concentration of species $k$ for the $j$-th of thermo-chemical state $[C_k]_j$---with $1 \le k \le N_{\text{sp}}$ the number of species in the model, and $1 \le j \le N_{\text{state}}$ the number of thermo-chemical states considered---there are two data-ordering choices.
+In the ``F'' (Fortran column-major) format, the concentrations of a single species $k$ over all thermo-chemical states are adjacent.
+In Fig.~\ref{F:mem} this would correspond to storing $[C_1]_1$ in index \num{0}, $[C_1]_2$ in index \num{1} and so on, with $N = N_{\text{state}}$, and $K = N_{\text{sp}}$.
+In the ``C'' (C row-major) format, the storage order is transposed with $[C_1]_1$ in index \num{0}, $[C_2]_1$ in index \num{1} etc., and $N = N_{\text{sp}}$, $K = N_{\text{state}}$.
 This ordering---along with the device (CPU, GPU, etc.) in question---have a large effect on the performance of SIMD\slash SIMT-vectorized algorithms.
@@
 In a \textit{shallow}-SIMD vectorization, each SIMD-lane evaluates chemical source terms for a different thermo-chemical state.
-If the data is stored in ``F''-order, the SIMD-lanes accessing the concentration of species $k$ for states $j, j+1\ldots$ will load sequential locations in memory, however the $k+1$th species concentration will be $N$ memory locations away increasing the likelihood of cache-misses on the CPU.
+If the data is stored in ``F''-order, the SIMD-lanes accessing the concentration of species $k$ for states $1, 2\ldots$ will load sequential locations in memory, however the $(k+1)$th species concentration will be $N_{\text{state}}$ memory locations away increasing the likelihood of cache-misses on the CPU.
 In a \textit{deep}-SIMD vectorization, a vector-unit cooperates to evaluate chemical source terms for a single thermo-chemical state $j$, thus SIMD-lanes loading species concentrations $k, k+1\ldots$ will access sequential memory locations if the data is stored in ``C''-order.
-Further, in the ``C''-ordering the furthest difference between any two concentrations in a single thermo-chemical state is at most $N_s$, with $N_s \ll N$ in most cases; this greatly improved data-locality increases the chances of a cache-hit on the CPU.
+Further, in the ``C''-ordering the furthest difference between any two concentrations in a single thermo-chemical state is at most $N_{\text{sp}}$, with $N_{\text{sp}} \ll N_{\text{state}}$ in most cases; this greatly improved data-locality increases the chances of a cache-hit on the CPU.
 However, a deep vectorization may result in SIMD-\textit{waste}---similar to thread-divergence in GPU-processing---caused by different SIMD-lanes executing different instructions (e.g.
@@
 Finally, in a SIMT-vectorization, threads execute the same instruction over multiple data in lock-step without use of specialized vectorization instructions.
-This is the paradigm used in GPU-processing, a more through review of which can be found in our previous work~\cite{CurtisGPU:2017}; one key-exception in this case is that the algorithm executed to evaluate the chemical source terms is identical for all thermo-chemical states, thus there is no thread-divergence for a shallow SIMT-vectorization.
+This is the paradigm used in GPU-processing, a more thorough review of which can be found in our previous work~\cite{CurtisGPU:2017}; one key-exception in this case is that the algorithm executed to evaluate the chemical source terms is identical for all thermo-chemical states, thus there is no thread-divergence for a shallow SIMT-vectorization.
 In Sec.~\ref{S:results} the effect of the data-ordering on the shallow-SIMD\slash SIMT vectorization performance will be studied.
@@
 \subsection{Previous works and goals of this study}
-Several recent works, e.g. \cite{CurtisGPU:2017} have investigated SIMT-based chemical kinetic integration on GPUs, however SIMD-based chemical kinetics have been less studied in comparison.
-\textcite{stone2016} implemented a linearly-implicit fourth-order stiff Rosenbrock solver in the OpenCL for various platforms including the CPU, GPU and the MIC.
+Several recent works, e.g. \cite{CurtisGPU:2017}, have investigated SIMT-based chemical kinetic integration on GPUs, however SIMD-based chemical kinetics have been less studied in comparison.
+\textcite{stone2016} implemented a linearly-implicit fourth-order stiff Rosenbrock solver in the OpenCL for various platforms including CPUs, GPUs, and MICs.
 The shallow-SIMD and SIMT-vectorization models were implemented in OpenCL and compared to an OpenMP baseline code that was deep SIMD-vectorized by simple compiler hints (a.k.a pragmas).
 The shallow-SIMD vectorization improved the integrator performance over the OpenMP baseline by \SIrange{2.5}{2.8}{$\times$} on the CPU and \SIrange{4.7}{4.9}{$\times$} on the MIC, while the GPU performance was \SIrange{1.4}{1.6}{$\times$} slower than the OpenMP baseline due to thread-divergence concerns.
-\textcite{kroshko2013efficient} implemented a shallow-SIMD vectorized third order stiff Rosenbrock integrator for atmospheric chemistry on a Cell Broadband Engine---a specially designed vector processor---finding a speedup of \SI{1.89}{$\times$} over a serial version of the same code, a \SI{94}{\percent} parallel efficiency.
+\textcite{kroshko2013efficient} implemented a shallow-SIMD vectorized third order stiff Rosenbrock integrator for atmospheric chemistry on a Cell Broadband Engine---a specially designed vector processor---finding a speedup of \SI{1.89}{$\times$} over a serial version of the same code.
 
@@
 \begin{equation}
-\frac{\text{d}[C_k]}{\text{d}t} = \sum_{i}^{N_R} \nu_{k,i} R_i c_i
+\dot{\omega}_k = \sum_{i=1}^{N_{\text{reac}}} \nu_{ki} R_i c_i
 \end{equation}
-where $\frac{\text{d}[C_k]}{\text{d}t}$, $N_R$, $\nu_{k,i}$, ${R_i}$, $c_i$ are the time rate of change of the concentration of species $k$, the number of reactions in the model, the net stoichiometric coefficient of species $k$ in reaction $i$, the net rate of progress of reaction $i$, and the pressure modification of reaction $i$ respectively.
+where $\dot{\omega}_k$, $N_{\text{reac}}$, $\nu_{k,i}$, ${R_i}$, $c_i$ are the time rate of change of the concentration of species $k$, the number of reactions in the model, the net stoichiometric coefficient of species $k$ in reaction $i$, the net rate of progress of reaction $i$, and the pressure modification of reaction $i$, respectively.
 For further detail on each term, the reader is referred to our previous work~\cite{Niemeyer:2016aa}; the code is capable of evaluating all modern reaction rate types (e.g.
-pressure-log, chebyshev, etc.).
+pressure-log, Chebyshev, etc.).
 
@@
 \begin{equation}
-\frac{\text{d}[T]}{\text{d}t} = -\frac{\sum_{k}^{N_S} H^{\degree}_k \frac{\text{d}[C_k]}{\text{d}t}}{\sum_{k}^{N_S} [C_k] c_{p_k}^{\degree}}
+\frac{\partial T}{\partial t} = -\frac{\sum_{k=1}^{N_{\text{sp}}} H^{\degree}_k \dot{\omega}_k}{\sum_{k=1}^{N_{\text{sp}}} [C_k] c_{p_k}^{\degree}}
 \end{equation}
-where $H^{\degree}_k$ and $c_{p_k}^{\degree}$ are the molar specific enthalpy and specific heat respectively.
+where $H^{\degree}_k$ and $c_{p_k}^{\degree}$ are the molar specific enthalpy and specific heat, respectively.
 Although the code is equally capable of a evaluating the temperature rate using a constant-volume assumption, we omit this here for brevity.
@@
 \section{Results and Discussion}
-\subsection{Validation}
+\subsection{Comparison to Cantera}
 The reaction rates of progress (ROP), species and temperature rates in this study are validated by comparison with Cantera~\cite{Cantera}, however special care must be taken due floating point arithmetic issues
 
-In computing the net ROP of reaction $i$ from the forward and reverse ROP: $R_{i} = R_{i}^{\prime} - R_{i}^{\prime\prime}$, precision can easily be lost as the net ROP may be---particularly near chemical equilibrium---many orders of magnitude smaller than the forward or reverse rates.
+In computing the net ROP of reaction $i$ for state $j$ from the forward and reverse ROP: $R_{ij} = R_{ij}^{\prime} - R_{ij}^{\prime\prime}$, precision can easily be lost as the net ROP may be---particularly near chemical equilibrium---many orders of magnitude smaller than the forward or reverse rates.
 A base-ten bound on loss of precision~\cite{goldberg1991every} can be found as:
 \begin{equation}
-\text{L}_{i,\text{min}} = 2^{-p_i} \le \left\lvert 1 - \frac{R_{i}^{\prime}}{R_{i}^{\prime\prime}} \right\rvert \le 2^{-q_i} = \text{L}_{i,\text{max}}}
+\text{L}_{ij,\text{lower}} = 2^{-p_i} \le \left\lvert 1 - \frac{R_{ij}^{\prime}}{R_{ij}^{\prime\prime}} \right\rvert \le 2^{-q_i} = \text{L}_{ij,\text{upper}}
 \label{e:bounds}
@@
 with $q_i \le p_i$.
-To obtain an estimate of the precision bounds, the forward and reverse ROP from Cantera are substituted into Eq.~\eqref{e:bounds}.
-Additionally a higher-precision number format (in this case, the Python Decimal class) is used in this computation to maintain accuracy.
-Finally, when determining the error in the net ROP between this code and Cantera, the bounds on the percent error attributable to precision loss, $\text{P}_{i,\text{err}}$ can be estimated as:
+To obtain an estimate of the precision bounds, the forward and reverse ROP from Cantera~\cite{Cantera} are substituted into Eq.~\eqref{e:bounds}.
+Additionally a higher-precision number format (in this case, a quad-precision floating point) is used in this computation to maintain accuracy.
+Finally, when determining the error in the net ROP between this code and Cantera~\cite{Cantera}, the bounds on the percent error attributable to precision loss, $\text{P}_{i,\text{err}}$ can be estimated as:
 \begin{equation}
-\text{P}_{i,\text{min}} = 100 * \frac{\text{L}_{i,\text{min}}}{\left\lvert\ R_{i,\text{CT}} - R_{i}\right\rvert} \le
-\text{P}_{i,\text{err}} \le 100 * \frac{\text{L}_{i,\text{max}}}{\left\lvert R_{i,\text{CT}} - R_{i} \right\rvert} = \text{P}_{i,\text{max}}
+\text{P}_{ij,\text{lower}} = 100 * \frac{\text{L}_{ij,\text{lower}}}{\left\lvert\ R_{ij,\text{CT}} - R_{ij}\right\rvert} \le
+\text{P}_{ij,\text{err}} \le 100 * \frac{\text{L}_{ij,\text{upper}}}{\left\lvert R_{ij,\text{CT}} - R_{ij} \right\rvert} = \text{P}_{ij,\text{upper}}
 \label{e:rel_bound}
 \end{equation}
-where the \text{CT} subscript indicates values from Cantera.
+where the \text{CT} subscript indicates values from Cantera~\cite{Cantera}.
+In the case that $R_{ij,\text{CT}} = R_{ij}$, the error of this net ROP is zero and the precision loss is ignored.
 
-For a more direct comparison, a relative error norm of a quantity $X_{i,j}$ over all states $j$, and indices $i$ was computed using the $L^{\infty}$ norm:
+For a more direct comparison, a relative error norm of a quantity $X_{ij}$ over all states $j$, and reactions $i$ was computed using the $L^{\infty}$ norm:
 \begin{equation}
- E_{X_{i}} = \left\lVert \frac{\left\lvert X_{i,j,\text{CT}} - X_{i,j}\right\rvert}{\num{e-10} + \num{e-6} * \left\lvert X_{i,j,\text{CT}} \right\rvert} \right\rVert_{i,j,\infty}
+ E_{X} = \left\lVert \frac{\left\lvert X_{ij,\text{CT}} - X_{ij}\right\rvert}{\num{e-10} + \num{e-6} * \left\lvert X_{i,j,\text{CT}} \right\rvert} \right\rVert_{i,j,\infty}
 \label{e:rel_err}
 \end{equation}
-and the lower precision bound for the net ROP error norm is set to $\text{P}_{\text{min}}^{*} = \text{P}_{i,\text{min}}$ for the $i, j$ value that results in the maximum error $E_{R_i}$.
+For the reaction $i$ and state $j$ that results in the maximum ROP error $E_{R}$, the lower precision bound is set to $\text{P}_{\text{lower}}^{*} = \text{P}_{i,\text{lower}}$.
 
@@
 \toprule
-\multicolumn{1}{l}{Model} & \multicolumn{1}{c}{$E_{R_{i}^{\prime}}$} & \multicolumn{1}{c}{$E_{R_{i}^{\prime\prime}}$} &\multicolumn{1}{c}{$E_{R_{i}}$} & \multicolumn{1}{c}{$\text{P}_{\text{min}}^{*}\si{\percent}$} & \multicolumn{1}{c}{$E_{\frac{\text{d}[C_k]}{\text{d}t}}$} & \multicolumn{1}{c}{$E_{\frac{\text{d}[T]}{\text{d}t}}$} \\
+\multicolumn{1}{l}{Model} & \multicolumn{1}{c}{$E_{R^{\prime}}$} & \multicolumn{1}{c}{$E_{R^{\prime\prime}}$} &\multicolumn{1}{c}{$E_{R}$} & \multicolumn{1}{c}{$\text{P}_{\text{lower}}^{*}\si{\percent}$} & \multicolumn{1}{c}{$E_{\dot{\omega}}$} & \multicolumn{1}{c}{$E_{\frac{\partial T}{\partial t}}$} \\
 \midrule
@@
 \multicolumn{1}{l}{USC-Mech II\cite{Wang:2007}}  & 1.07e-7 & 1.37e-7 & 2.39e0 & 4.12e02 & 7.92e00 & 3.22e03 \\
-\multicolumn{1}{l}{\ce{IC5H11OH}\cite{Sarathy:2013jr}} & -1 &  -1 & -1 & -1 & -1 & -1 \\
+\multicolumn{1}{l}{\ce{iC5H11OH}\cite{Sarathy:2013jr}} & -1 &  -1 & -1 & -1 & -1 & -1 \\
 \bottomrule
@@
 
-In Table~\ref{T:error}, we see the results of this code as compared to Cantera on a library of partially stirred reaction conditions (PaSR) described in our previous works~\cite{CurtisGPU:2017,Niemeyer:2016aa}.
+In Table~\ref{T:error}, we see the results of this code as compared to Cantera~\cite{Cantera} on a library of partially stirred reaction conditions (PaSR) described in our previous works~\cite{CurtisGPU:2017,Niemeyer:2016aa}.
 Very close agreement is observed for the forward and reverse ROP for all models, however the net ROP error norm is \numrange{7}{8} orders of magnitude larger.
-We see here that the minimum estimated loss of precision is \SIrange{4}{40}{$\times$} the error in net ROP, explaining this jump.
+We see here that the estimated lower bound on loss of precision is \SIrange{4}{40}{$\times$} the error in net ROP, explaining the large increase in the net ROP error norm as compared to either the forward or backward ROP error norms.
 The species rate error norm is similar in magnitude to that of the net ROP, however the temperature rate error again increases as it depends directly on the species rates.
-We note that the above discussion does not imply that computation of the net ROP will cause large error in chemical kinetic integration---either in this code or Cantera---as this loss of precision only occurs when the forward and reverse ROP are nearly equal, implying the reaction is in near-equilibrium.
+We note that the above discussion does not imply that computation of the net ROP will cause large error in chemical kinetic integration---either in this code or Cantera~\cite{Cantera}---as this loss of precision only occurs when the forward and reverse ROP are nearly equal, implying the reaction is in near-equilibrium.
 
-\subsection{Results}
+\subsection{Performance}
 \label{S:results}
@@
 Significant speedups are seen with larger vector-widths for the smaller models, however as the model size increases data must be loaded from increasingly distant memory locations and the performance of all three vector-widths becomes similar.
-Finally in Fig.~\ref{F:SIMDscale} the strong parallel scaling of the chemical source term evaluation is examined for all four models with a ``C''-ordered shallow SIMD-vectorization and vector-width of 16:
-Interestingly the efficiency for the hydrogen model decreases fastest, while the larger models have similar efficiencies; the isopentanol model has the highest efficiency in all cases.
+Finally, the strong parallel scaling efficiency:
+\begin{equation}
+S = \frac{t_1}{N t_N}
+\end{equation}
+where $t_1$ is the mean wall-clock runtime of the chemical source terms evaluation on one core and $t_N$ the mean runtime on $N$ cores, is examined in Fig~.\ref{F:SIMDscale} for all four models with a ``C''-ordered shallow SIMD-vectorization and vector-width of 16.
+Interestingly the efficiency for the hydrogen model decreases fastest, while the larger models have nearly identical for more than two cores; the isopentanol model has the highest efficiency in all cases.
 This suggests that evaluation of the chemical-source terms for a single thermo-chemical state for the smaller models may not saturate the processing power of the CPU, some efficiency may be lost due to context switching within OpenCL itself.
-Additionally, OpenCL allows assigning of work to devices by shared memory resources e.g., a L3 cache, which was not pursued in this study for simplicity; this should be investigated in future work to determine the effect on scaling efficiency.
+Additionally, OpenCL allows assigning of work to devices by shared memory resources, e.g., a L3 cache, which was not pursued in this study for simplicity; this should be investigated in future work to determine the effect on scaling efficiency.
 
 \section{Conclusions}
-In this work, a shallow SIMD-vectorized chemical source term evaluation OpenCL code generator was developed and validated against Cantera~\cite{Cantera} for a wide range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr}.
+In this work, automatically generated OpenCL codes for shallow SIMD-vectorized chemical source term evaluation were developed and validated against Cantera~\cite{Cantera} for a wide range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr}.
 Significant speedups of up to \SIrange{1.88}{2.42}{$\times$} over a baseline SIMT-vectorized code were observed.
@@
 Further, this study suggests that a deep SIMD-vectorized code---currently under development---may see even greater accelerations due to expected increases in cache-hit rates.
-Larger vector-widths were found to provide accelerations for the smaller models studied, but performance was similar over all vector-widths tested for the largest models tested.
+Larger vector-widths were found to provide accelerations for the smaller models studied, but the speedup was similar over all vector-widths tested for the largest models investigated.
 Finally, the strong parallel scaling efficiency was examined for the shallow vectorized code; the larger models exhibited higher scaling efficiency, possibly due to failure of the \ce{H2}\slash\ce{CO} model to saturate the CPU throughput.
-Future extensions of this work will include development of a SIMD\slash SIMT-accelerated sparse-analytical Jacobian code targeted at CPUs, GPUs and MICs.
+Future extensions of this work will include development of a SIMD\slash SIMT-accelerated sparse-analytical Jacobian code targeted at CPUs, GPUs, and MICs.
 
 \section{Acknowledgements}
-This research was funded by \ldots
+This research was funded by the National Science Foundation under grant ACI-1534688.
 

### 2017-02-23 scripts/general_plotting.py
plot updates, py3 compat

@@ -29,7 +29,7 @@ def process_data(plotdata, plot, reacs_as_x=True,
         zs = np.std(ys)
         if plot == 'runtime':
             ys = np.array([getattr(data, plot) / data.num_conditions for data in run.rundata])
-            zs /= data.num_conditions
+            zs /= run.rundata[0].num_conditions
         err_vals.append(zs)
         y_vals.append(np.mean(ys))
     return x_vals, y_vals, err_vals

### 2017-02-23 scripts/perf_plotter.py

@@ -50,7 +50,7 @@ def plotter(plot_name='', show=True, plot_reacs=True, norm=True,
         if filters[f] is None:
             continue
         assert any(data[x] for x in data), 'No data matching all filters'
-        if f in data[data.keys()[0]][0]._asdict():
+        if f in data[list(data.keys())[0]][0]._asdict():
             for mech in data:
                 data[mech] = [x for x in data[mech] if __compare(x, f, filters[f])]
     #compute data
@@ -212,7 +212,7 @@ def plotter(plot_name='', show=True, plot_reacs=True, norm=True,
     if norm:
         ylabel = 'Speedup'
         if plot_cores:
-            ylabel = 'Strong Parallel Scaling'
+            ylabel = 'Parallel Scaling Efficiency'
     if plot_cores:
         xlabel = 'Number of Cores'
 

### 2017-02-23 scripts/simd_and_order_comp.py

@@ -26,7 +26,7 @@ def marker_func(name):
         retval.append(SIMD_color)
     return tuple(retval)
 
-labels = ['SIMT', 'SIMD', 'C-order', 'F-order']
+labels = ['SIMT', 'SIMD', 'C-order (Filled)', 'F-order (Hollow)']
 handles = [mlines.Line2D([], [], color=SIMT_color, marker=SIMT_sym, linestyle='',
                           markersize=ps.marker_style['size'], label='SIMT',
                           markerfacecoloralt='none',
@@ -35,10 +35,8 @@ handles = [mlines.Line2D([], [], color=SIMT_color, marker=SIMT_sym, linestyle=''
                           markersize=ps.marker_style['size'], label='SIMD',
                           markerfacecoloralt='none',
                           fillstyle='left'),
-           mlines.Line2D([], [], linestyle='', marker=(4, 0, 45), markersize=ps.marker_style['size'],
-                          label='C-order', markerfacecolor='k', markeredgecolor='k'),
-           mlines.Line2D([], [], linestyle='', marker=(4, 0, 45), markersize=ps.marker_style['size'],
-                          label='F-order', markerfacecolor='none', markeredgecolor='k')]
+           mlines.Line2D([], [], linestyle='', marker=None, label='C-order (Filled)'),
+           mlines.Line2D([], [], linestyle='', marker=None, label='F-order (Hollow)')]
 legend = (handles, labels)
 
 xval, yval, zval, label = pp.plotter('../SIMD_SIMT_Order_comparison.pdf', show=False,
@@ -55,12 +53,12 @@ C_SIMD = next(i for i in C_order if i != C_SIMT)
 F_SIMT = next(i for i in F_order if 'SIMT' in label[i])
 F_SIMD = next(i for i in F_order if i != F_SIMT)
 
-print 'SIMT vs SIMD'
+print('SIMT vs SIMD')
 for i in range(len(yval[0])):
-    print i, 'C', yval[C_SIMT][i] / yval[C_SIMD][i]
-    print i, 'F', yval[F_SIMT][i] / yval[F_SIMD][i]
+    print(i, 'C', yval[C_SIMT][i] / yval[C_SIMD][i])
+    print(i, 'F', yval[F_SIMT][i] / yval[F_SIMD][i])
 
-print 'C vs F'
+print('C vs F')
[…]

### 2017-02-23 scripts/simd_and_order_comp.py
fix markers

@@ -28,13 +28,9 @@ def marker_func(name):
 
 labels = ['SIMT', 'SIMD', 'C-order (Filled)', 'F-order (Hollow)']
 handles = [mlines.Line2D([], [], color=SIMT_color, marker=SIMT_sym, linestyle='',
-                          markersize=ps.marker_style['size'], label='SIMT',
-                          markerfacecoloralt='none',
-                          fillstyle='left'),
+                          markersize=ps.marker_style['size'], label='SIMT'),
            mlines.Line2D([], [], color=SIMD_color, marker=SIMD_sym, linestyle='',
-                          markersize=ps.marker_style['size'], label='SIMD',
-                          markerfacecoloralt='none',
-                          fillstyle='left'),
+                          markersize=ps.marker_style['size'], label='SIMD'),
            mlines.Line2D([], [], linestyle='', marker=None, label='C-order (Filled)'),
            mlines.Line2D([], [], linestyle='', marker=None, label='F-order (Hollow)')]
 legend = (handles, labels)

### 2017-02-24 scripts/error.py
updates moving to laptop

@@ -17,7 +17,7 @@ for mech in mechs:
             name = line[:line.index(':')].strip()
             err_arr = np.array([float(x) for x in line[line.index(':') + 1:].split(',') if x.strip()])
             if ('rop_fwd' in name or 'rop_rev' in name) and 'linf' in name and np.any(err_arr > 1):
-                print file, np.where(err_arr > 1)
+                print(file, np.where(err_arr > 1))
             if name in err_dicts[mech_name]:
                 try:
                     assert np.allclose(err_dicts[mech_name][name], err_arr)

### 2017-02-24 paper.tex
updates moving to laptop
update error stats

@@
 \multicolumn{1}{l}{USC-Mech II\cite{Wang:2007}}  & 1.07e-7 & 1.37e-7 & 2.39e0 & 4.12e02 & 7.92e00 & 3.22e03 \\
-\multicolumn{1}{l}{\ce{iC5H11OH}\cite{Sarathy:2013jr}} & -1 &  -1 & -1 & -1 & -1 & -1 \\
+\multicolumn{1}{l}{\ce{iC5H11OH}\cite{Sarathy:2013jr}} & 6.80e-8 &  1.48e-7 & 5.26e-1 & 4.30e-1 & 2.26e00 & 6.10e+01 \\
 \bottomrule
@@
 Error statistics are based on the infinity-norm of the relative error detailed in Eq.~\eqref{e:rel_err} for each quantity.
-\todo[inline]{i-pentanol validation still running}
 }
@@
 In Table~\ref{T:error}, we see the results of this code as compared to Cantera~\cite{Cantera} on a library of partially stirred reaction conditions (PaSR) described in our previous works~\cite{CurtisGPU:2017,Niemeyer:2016aa}.
-Very close agreement is observed for the forward and reverse ROP for all models, however the net ROP error norm is \numrange{7}{8} orders of magnitude larger.
+Very close agreement is observed for the forward and reverse ROP for all models, however the net ROP error norm is \numrange{6}{8} orders of magnitude larger.
 We see here that the estimated lower bound on loss of precision is \SIrange{4}{40}{$\times$} the error in net ROP, explaining the large increase in the net ROP error norm as compared to either the forward or backward ROP error norms.

### 2017-02-24 scripts/error.py

@@ -44,7 +44,7 @@ for mech in err_dicts:
             maxv = np.linalg.norm(err_vals, ord=np.inf)
             maxind = np.where(err_dicts[mech][name] == maxv)[0][0]
             print(name, format(maxv))
-            print('precision_max', format(err_dicts[mech]['rop_net_precmax'][maxind]))
-            print('precision_min', format(err_dicts[mech]['rop_net_precmin'][maxind]))
+            print('precision_max', format(err_dicts[mech]['rop_net_precmax - linf'][maxind]))
+            print('precision_min', format(err_dicts[mech]['rop_net_precmin - linf'][maxind]))
         elif 'prec' not in name:
             print(name, format(np.linalg.norm(err_vals, ord=np.inf)))

### 2017-02-25 paper.tex
larger plots
minor updates

@@
 If the data is stored in ``F''-order, the SIMD-lanes accessing the concentration of species $k$ for states $1, 2\ldots$ will load sequential locations in memory, however the $(k+1)$th species concentration will be $N_{\text{state}}$ memory locations away increasing the likelihood of cache-misses on the CPU.
-In a \textit{deep}-SIMD vectorization, a vector-unit cooperates to evaluate chemical source terms for a single thermo-chemical state $j$, thus SIMD-lanes loading species concentrations $k, k+1\ldots$ will access sequential memory locations if the data is stored in ``C''-order.
+In a \textit{deep}-SIMD vectorization, a vector-unit cooperates to evaluate chemical source terms for a single thermo-chemical state $j$, thus SIMD-lanes loading species concentrations $k, k+1\ldots$ for e.g., state 1, will access sequential memory locations if the data is stored in ``C''-order.
 Further, in the ``C''-ordering the furthest difference between any two concentrations in a single thermo-chemical state is at most $N_{\text{sp}}$, with $N_{\text{sp}} \ll N_{\text{state}}$ in most cases; this greatly improved data-locality increases the chances of a cache-hit on the CPU.
@@
 \subsection{Previous works and goals of this study}
-Several recent works, e.g. \cite{CurtisGPU:2017}, have investigated SIMT-based chemical kinetic integration on GPUs, however SIMD-based chemical kinetics have been less studied in comparison.
+Several recent works, e.g.~\cite{CurtisGPU:2017}, have investigated SIMT-based chemical kinetic integration on GPUs, however SIMD-based chemical kinetics have been less studied in comparison.
 \textcite{stone2016} implemented a linearly-implicit fourth-order stiff Rosenbrock solver in the OpenCL for various platforms including CPUs, GPUs, and MICs.
@@
 \begin{equation}
-\text{L}_{ij,\text{lower}} = 2^{-p_i} \le \left\lvert 1 - \frac{R_{ij}^{\prime}}{R_{ij}^{\prime\prime}} \right\rvert \le 2^{-q_i} = \text{L}_{ij,\text{upper}}
+\text{L}_{ij,\text{lower}} = 2^{-p_{ij}} \le \left\lvert 1 - \frac{R_{ij}^{\prime}}{R_{ij}^{\prime\prime}} \right\rvert \le 2^{-q_{ij}} = \text{L}_{ij,\text{upper}}
 \label{e:bounds}
 \end{equation}
-with $q_i \le p_i$.
+with $q_{ij} \le p_{ij}$.
 To obtain an estimate of the precision bounds, the forward and reverse ROP from Cantera~\cite{Cantera} are substituted into Eq.~\eqref{e:bounds}.
 Additionally a higher-precision number format (in this case, a quad-precision floating point) is used in this computation to maintain accuracy.
-Finally, when determining the error in the net ROP between this code and Cantera~\cite{Cantera}, the bounds on the percent error attributable to precision loss, $\text{P}_{i,\text{err}}$ can be estimated as:
+Finally, when determining the error in the net ROP between this code and Cantera~\cite{Cantera}, the bounds on the percent error attributable to precision loss, $\text{P}_{ij,\text{err}}$ can be estimated as:
 \begin{equation}
@@
 \begin{equation}
- E_{X} = \left\lVert \frac{\left\lvert X_{ij,\text{CT}} - X_{ij}\right\rvert}{\num{e-10} + \num{e-6} * \left\lvert X_{i,j,\text{CT}} \right\rvert} \right\rVert_{i,j,\infty}
+ E_{X} = \left\lVert \frac{\left\lvert X_{ij,\text{CT}} - X_{ij}\right\rvert}{\num{e-10} + \num{e-6} * \left\lvert X_{ij,\text{CT}} \right\rvert} \right\rVert_{i,j,\infty}
 \label{e:rel_err}
@@
 We note that the above discussion does not imply that computation of the net ROP will cause large error in chemical kinetic integration---either in this code or Cantera~\cite{Cantera}---as this loss of precision only occurs when the forward and reverse ROP are nearly equal, implying the reaction is in near-equilibrium.
+\todo{update for new valid}
 
@@
     \includegraphics[width=\textwidth]{SIMD_scaling}
-    \caption{Strong parallel scaling of a ``C''-ordered shallow-SIMD vectorization using a vector-width of 16.}
+    \caption{Strong parallel scaling efficiency of a ``C''-ordered shallow-SIMD vectorization using a vector-width of 16.}
     \label{F:SIMDscale}

### 2017-02-25 paper.tex
clarify loading examples with Nv

@@
 A single processing element (e.g.
-a CPU core) contains a vector unit with $N$ lanes ($\text{L}_0,\ldots \text{L}_N$).
+a CPU core) contains a vector unit with $N_v$ lanes ($\text{L}_0,\ldots \text{L}_{N_v}$).
 The vector unit executes a single instruction concurrently on multiple data.}
@@
 In a \textit{shallow}-SIMD vectorization, each SIMD-lane evaluates chemical source terms for a different thermo-chemical state.
-If the data is stored in ``F''-order, the SIMD-lanes accessing the concentration of species $k$ for states $1, 2\ldots$ will load sequential locations in memory, however the $(k+1)$th species concentration will be $N_{\text{state}}$ memory locations away increasing the likelihood of cache-misses on the CPU.
-In a \textit{deep}-SIMD vectorization, a vector-unit cooperates to evaluate chemical source terms for a single thermo-chemical state $j$, thus SIMD-lanes loading species concentrations $k, k+1\ldots$ for e.g., state 1, will access sequential memory locations if the data is stored in ``C''-order.
+If the data is stored in ``F''-order, the SIMD-lanes accessing the concentration of species $k$ for e.g., states $1, 2,\ldots N_{v}$---the number of vector lanes---will load sequential locations in memory, however the $(k+1)$th species concentration will be $N_{\text{state}}$ memory locations away increasing the likelihood of cache-misses on the CPU.
+In a \textit{deep}-SIMD vectorization, a vector-unit cooperates to evaluate chemical source terms for a single thermo-chemical state, thus SIMD-lanes loading species concentrations $k, k+1\ldots k + N_v$ for e.g., state 1, will access sequential memory locations if the data is stored in ``C''-order.
 Further, in the ``C''-ordering the furthest difference between any two concentrations in a single thermo-chemical state is at most $N_{\text{sp}}$, with $N_{\text{sp}} \ll N_{\text{state}}$ in most cases; this greatly improved data-locality increases the chances of a cache-hit on the CPU.

### 2017-02-25 paper.tex
more minor edits

@@
 \end{equation}
-where $\dot{\omega}_k$, $N_{\text{reac}}$, $\nu_{k,i}$, ${R_i}$, $c_i$ are the time rate of change of the concentration of species $k$, the number of reactions in the model, the net stoichiometric coefficient of species $k$ in reaction $i$, the net rate of progress of reaction $i$, and the pressure modification of reaction $i$, respectively.
+where $\dot{\omega}_k$, $N_{\text{reac}}$, $\nu_{ki}$, ${R_i}$, $c_i$ are the time rate of change of the concentration of species $k$, the number of reactions in the model, the net stoichiometric coefficient of species $k$ in reaction $i$, the net rate of progress of reaction $i$, and the pressure modification of reaction $i$, respectively.
 For further detail on each term, the reader is referred to our previous work~\cite{Niemeyer:2016aa}; the code is capable of evaluating all modern reaction rate types (e.g.
@@
 \end{equation}
-For the reaction $i$ and state $j$ that results in the maximum ROP error $E_{R}$, the lower precision bound is set to $\text{P}_{\text{lower}}^{*} = \text{P}_{i,\text{lower}}$.
+For the reaction $i$ and state $j$ that results in the maximum ROP error $E_{R}$, the lower precision bound is set to $\text{P}_{\text{lower}}^{*} = \text{P}_{ij,\text{lower}}$.
 
@@
 Very close agreement is observed for the forward and reverse ROP for all models, however the net ROP error norm is \numrange{6}{8} orders of magnitude larger.
-We see here that the estimated lower bound on loss of precision is \SIrange{4}{40}{$\times$} the error in net ROP, explaining the large increase in the net ROP error norm as compared to either the forward or backward ROP error norms.
+We see here that the estimated lower bound on loss of precision is \SIrange{4}{40}{$\times$} the error in net ROP for all but the \ce{iC5H11OH} model, explaining the large increase in the net ROP error norm as compared to either the forward or backward ROP error norms.
 The species rate error norm is similar in magnitude to that of the net ROP, however the temperature rate error again increases as it depends directly on the species rates.
@@
 Significant speedups are seen with larger vector-widths for the smaller models, however as the model size increases data must be loaded from increasingly distant memory locations and the performance of all three vector-widths becomes similar.
-Finally, the strong parallel scaling efficiency:
+
+Finally, the strong parallel scaling efficiency is defined as:
 \begin{equation}
@@
 \end{equation}
-where $t_1$ is the mean wall-clock runtime of the chemical source terms evaluation on one core and $t_N$ the mean runtime on $N$ cores, is examined in Fig~.\ref{F:SIMDscale} for all four models with a ``C''-ordered shallow SIMD-vectorization and vector-width of 16.
-Interestingly the efficiency for the hydrogen model decreases fastest, while the larger models have nearly identical for more than two cores; the isopentanol model has the highest efficiency in all cases.
+where $t_1$ is the mean wall-clock runtime of chemical source terms evaluation on one core and $t_N$ the mean runtime on $N$ cores.
+Fig.~\ref{F:SIMDscale} examines the strong parallel scaling efficiency for all four models with a ``C''-ordered shallow SIMD-vectorization and vector-width of 16.
+Interestingly the efficiency of the hydrogen model decreases the fastest while the larger models have significantly larger scaling efficiencies for more than 4 cores.
+Additionally, it appears that for a fixed number of cores, the scaling efficiency increases with model size.
 This suggests that evaluation of the chemical-source terms for a single thermo-chemical state for the smaller models may not saturate the processing power of the CPU, some efficiency may be lost due to context switching within OpenCL itself.

### 2017-02-25 paper.tex
update #'s for updated perf

@@
 A code generation platform for single-instruction, multiple-data (SIMD) vectorized chemical source term evaluation was developed and validated against Cantera for a wide range of chemical kinetic models.
-Speedups of up to \SIrange{1.88}{2.42}{$\times$} were observed over a baseline single-instruction, multiple-thread (SIMT) parallelized code.
-Additionally, use of a row-major data layout was found to increase both the SIMD and SIMT performance by up to \SIrange{1.96}{2.38}{$\times$} over a column-major data layout.
+Speedups of up to \SIrange{1.90}{2.45}{$\times$} were observed over a baseline single-instruction, multiple-thread (SIMT) parallelized code.
+Additionally, use of a row-major data layout was found to increase both the SIMD and SIMT performance by up to \SIrange{2.00}{2.38}{$\times$} over a column-major data layout.
 Larger vector-widths were found to produce speedups up to \SI{1.8}{$\times$} as compared to a small vector-width for smaller models, e.g.
@@
 In Fig.~\ref{F:SIMDComp}, the performance of the shallow-SIMD and SIMT vectorization are compared on a single CPU core for both ``C'' and ``F'' memory formats.
-The ``C''-ordered SIMD\slash SIMT vectorizations show a speedup over the equivalent ``F''-ordered code by \SIrange{1.36}{1.96}{$\times$} and \SIrange{1.60}{2.38}{$\times$} respectively.
+The ``C''-ordered SIMD\slash SIMT vectorizations show a speedup over the equivalent ``F''-ordered code by \SIrange{1.46}{2.00}{$\times$} and \SIrange{1.61}{2.38}{$\times$} respectively.
 This demonstrates that enhanced data-locality of the ``C''-ordering for CPU caching---discussed in Sec.~\ref{S:data}---is key to high performance on the CPU, and suggests that a deep vectorization would see even greater accelerations.
-The shallow vectorization is consistently faster than the SIMT vectorization, by \SIrange{1.72}{1.88}{$\times$} and \SIrange{2.10}{2.42}{$\times$} for the ``C'' and ``F'' memory formats respectively.
+The shallow vectorization is consistently faster than the SIMT vectorization, by \SIrange{1.73}{1.90}{$\times$} and \SIrange{2.00}{2.45}{$\times$} for the ``C'' and ``F'' memory formats respectively.
 This is less than the theoretical maximum speedup on this CPU, which can process \SI{256}{\bit} AVX instructions (i.e. four double-precision operations) concurrently.
@@
 In this work, automatically generated OpenCL codes for shallow SIMD-vectorized chemical source term evaluation were developed and validated against Cantera~\cite{Cantera} for a wide range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr}.
-Significant speedups of up to \SIrange{1.88}{2.42}{$\times$} over a baseline SIMT-vectorized code were observed.
+Significant speedups of up to \SIrange{1.90}{2.45}{$\times$} over a baseline SIMT-vectorized code were observed.
 Two data-ordering schemes were investigated, showing a clear performance benefit for use of the ``C'' (row-major) ordering for the shallow vectorized code.

### 2017-02-26 scripts/error.py
do normalization in post processing, find rop component instead of precision

@@ -4,34 +4,47 @@ script_dir = os.path.dirname(os.path.abspath(__file__))
 error_path = os.path.join(script_dir, 'error_checking')
 mechs = [os.path.join(error_path, x) for x in os.listdir(error_path) if os.path.isdir(os.path.join(error_path, x))]
 
+rtol = 1e-6
+atol = 1e-10
+
 err_dicts = {}
 for mech in mechs:
     mech_name = os.path.basename(os.path.normpath(mech))
     #get file list
-    files = [os.path.join(mech,x) for x in os.listdir(mech) if os.path.isfile(os.path.join(mech,x))]
+    files = [os.path.join(mech,x) for x in os.listdir(mech) if os.path.isfile(os.path.join(mech,x)) if x.endswith('.npz')]
     err_dicts[mech_name] = {}
     for file in files:
-        with open(file, 'r') as f:
-            lines = [x for x in f.read().split('\n') if x.strip()]
-        for line in lines:
-            name = line[:line.index(':')].strip()
-            err_arr = np.array([float(x) for x in line[line.index(':') + 1:].split(',') if x.strip()])
-            if ('rop_fwd' in name or 'rop_rev' in name) and 'linf' in name and np.any(err_arr > 1):
-                print(file, np.where(err_arr > 1))
+        arrs = np.load(file)
+        for name in arrs:
+            if 'value' in name or 'component' in name:
+                continue
+            errs = arrs[name]
+            values = arrs[name + '_value']
+            errs = errs / (atol + rtol * np.abs(values))
+
+            precs = None
+            if 'rop_net' in name:
+                #calculate the precision norms
+                precs = 100 * arrs['rop_component'] / arrs[name]
+
             if name in err_dicts[mech_name]:
-                try:
-                    assert np.allclose(err_dicts[mech_name][name], err_arr)
-                    err_dicts[mech_name][name] = err_arr
-                except AssertionError:
-                    err_dicts[mech_name][name] = np.maximum(err_dicts[mech_name][name], err_arr)
+                err_dicts[mech_name][name] = np.maximum(err_dicts[mech_name][name], errs)
+                if 'rop_net' in name:
+                    update_locs = np.where(err_dicts[mech_name][name] == errs)
+                    #update the precision norms at these locations
+                    err_dicts[mech_name]['rop_component'][update_locs] = precs[update_locs]
             else:
[…]

### 2017-02-26 paper.tex
update the validation section

@@
 \subsection{Comparison to Cantera}
-The reaction rates of progress (ROP), species and temperature rates in this study are validated by comparison with Cantera~\cite{Cantera}, however special care must be taken due floating point arithmetic issues
+The reaction rates of progress (ROP), species and temperature rates in this study are validated by comparison with Cantera~\cite{Cantera}, however special care must be taken due floating point arithmetic issues.
 
-In computing the net ROP of reaction $i$ for state $j$ from the forward and reverse ROP: $R_{ij} = R_{ij}^{\prime} - R_{ij}^{\prime\prime}$, precision can easily be lost as the net ROP may be---particularly near chemical equilibrium---many orders of magnitude smaller than the forward or reverse rates.
-A base-ten bound on loss of precision~\cite{goldberg1991every} can be found as:
-\begin{equation}
-\text{L}_{ij,\text{lower}} = 2^{-p_{ij}} \le \left\lvert 1 - \frac{R_{ij}^{\prime}}{R_{ij}^{\prime\prime}} \right\rvert \le 2^{-q_{ij}} = \text{L}_{ij,\text{upper}}
-\label{e:bounds}
-\end{equation}
-with $q_{ij} \le p_{ij}$.
-To obtain an estimate of the precision bounds, the forward and reverse ROP from Cantera~\cite{Cantera} are substituted into Eq.~\eqref{e:bounds}.
-Additionally a higher-precision number format (in this case, a quad-precision floating point) is used in this computation to maintain accuracy.
-Finally, when determining the error in the net ROP between this code and Cantera~\cite{Cantera}, the bounds on the percent error attributable to precision loss, $\text{P}_{ij,\text{err}}$ can be estimated as:
-\begin{equation}
-\text{P}_{ij,\text{lower}} = 100 * \frac{\text{L}_{ij,\text{lower}}}{\left\lvert\ R_{ij,\text{CT}} - R_{ij}\right\rvert} \le
-\text{P}_{ij,\text{err}} \le 100 * \frac{\text{L}_{ij,\text{upper}}}{\left\lvert R_{ij,\text{CT}} - R_{ij} \right\rvert} = \text{P}_{ij,\text{upper}}
-\label{e:rel_bound}
-\end{equation}
-where the \text{CT} subscript indicates values from Cantera~\cite{Cantera}.
-In the case that $R_{ij,\text{CT}} = R_{ij}$, the error of this net ROP is zero and the precision loss is ignored.
-
-For a more direct comparison, a relative error norm of a quantity $X_{ij}$ over all states $j$, and reactions $i$ was computed using the $L^{\infty}$ norm:
+For a direct comparison, a relative error norm of a quantity $X_{ij}$ over all states $j$, and reactions (or species) $i$ was computed using the $L^{\infty}$ norm:
 \begin{equation}
@@
 \end{equation}
-For the reaction $i$ and state $j$ that results in the maximum ROP error $E_{R}$, the lower precision bound is set to $\text{P}_{\text{lower}}^{*} = \text{P}_{ij,\text{lower}}$.
+where the \text{CT} subscript indicates values from Cantera~\cite{Cantera}
+
+However, in computing the net ROP of reaction $i$ for state $j$ from the forward and reverse ROP: $R_{ij} = R_{ij}^{\prime} - R_{ij}^{\prime\prime}$, accuracy can easily be lost as the net ROP may be---particularly near chemical equilibrium---many orders of magnitude smaller than the forward or reverse rates.
+To quantify this phenomena, the error in forward ROP is first defined as:
+\begin{equation}
+\varepsilon^{\prime}_{ij} = \left\lvert R_{ij}^{\prime} - R_{ij,\text{CT}}^{\prime} \right\rvert
+\end{equation}
+and the error in reverse ROP $\varepsilon^{\prime\prime}_{ij}$ can be defined analogously.
+Finally, for the reaction $i^{*}$ and state $j^{*}$ that result in the largest error in net ROP, i.e.
+$E_{R}$, an estimate of the percent error attributable to floating point error accumulation from the forward and reverse ROPs can be obtained as:
+\begin{equation}
+E_{\varepsilon} = 100\times\frac{\max(\varepsilon^{\prime}_{i^{*}j^{*}}\text{, }\varepsilon^{\prime\prime}_{i^{*}j^{*}})}{R_{i^{*}j^{*},\text{CT}}} \si{\percent}
+\end{equation}
+If this estimate is near \SI{100}{\percent}, it indicates that the maximum error in forward or reverse ROPs are of a similar magnitude to the net ROP, hence the error in net ROP would be expected to be large in this case.
 
@@
 \toprule
-\multicolumn{1}{l}{Model} & \multicolumn{1}{c}{$E_{R^{\prime}}$} & \multicolumn{1}{c}{$E_{R^{\prime\prime}}$} &\multicolumn{1}{c}{$E_{R}$} & \multicolumn{1}{c}{$\text{P}_{\text{lower}}^{*}\si{\percent}$} & \multicolumn{1}{c}{$E_{\dot{\omega}}$} & \multicolumn{1}{c}{$E_{\frac{\partial T}{\partial t}}$} \\
+\multicolumn{1}{l}{Model} & \multicolumn{1}{c}{$E_{R^{\prime}}$} & \multicolumn{1}{c}{$E_{R^{\prime\prime}}$} &\multicolumn{1}{c}{$E_{R}$} & \multicolumn{1}{c}{$E_{\varepsilon}\si{\percent}$} & \multicolumn{1}{c}{$E_{\dot{\omega}}$} & \multicolumn{1}{c}{$E_{\frac{\partial T}{\partial t}}$} \\
 \midrule
-\multicolumn{1}{l}{\ce{H2}\slash~\ce{CO}\cite{Burke:2011fh}} & 1.20e-8 & 6.33e-8 & 1.32e1 & 4.66e6 & 2.37e1 & 2.12e5 \\
-\multicolumn{1}{l}{GRI-Mech.~3.0\cite{smith_gri-mech_30}}  & 3.07e-8 & 5.90e-8 & 1.21e0 & 2.67e04 & 2.84e0 & 2.64e4 \\
-\multicolumn{1}{l}{USC-Mech II\cite{Wang:2007}}  & 1.07e-7 & 1.37e-7 & 2.39e0 & 4.12e02 & 7.92e00 & 3.22e03 \\
-\multicolumn{1}{l}{\ce{iC5H11OH}\cite{Sarathy:2013jr}} & 6.80e-8 &  1.48e-7 & 5.26e-1 & 4.30e-1 & 2.26e00 & 6.10e+01 \\
+\multicolumn{1}{l}{\ce{H2}\slash~\ce{CO}\cite{Burke:2011fh}} & 1.20e-8 & 6.33e-8 & 1.32e1 & 1.08e2 & 2.37e1 & 2.12e5 \\
+\multicolumn{1}{l}{GRI-Mech.~3.0\cite{smith_gri-mech_30}}  & 3.07e-8 & 5.90e-8 & 1.21e0 & 1.42e2 & 2.84e0 & 2.64e4 \\
+\multicolumn{1}{l}{USC-Mech II\cite{Wang:2007}}  & 1.07e-7 & 1.37e-7 & 2.39e0 & 4.12e02 & 1.44e2 & 3.22e03 \\
+\multicolumn{1}{l}{\ce{iC5H11OH}\cite{Sarathy:2013jr}} & 6.80e-8 &  1.48e-7 & 5.26e-1 & 1.59e2 & 2.26e00 & 6.10e+01 \\
 \bottomrule
@@
 In Table~\ref{T:error}, we see the results of this code as compared to Cantera~\cite{Cantera} on a library of partially stirred reaction conditions (PaSR) described in our previous works~\cite{CurtisGPU:2017,Niemeyer:2016aa}.
-Very close agreement is observed for the forward and reverse ROP for all models, however the net ROP error norm is \numrange{6}{8} orders of magnitude larger.
-We see here that the estimated lower bound on loss of precision is \SIrange{4}{40}{$\times$} the error in net ROP for all but the \ce{iC5H11OH} model, explaining the large increase in the net ROP error norm as compared to either the forward or backward ROP error norms.
+Very close agreement is observed for the forward and reverse ROP for all models, however the net ROP error norm is \numrange{6}{9} orders of magnitude larger.
+We see here that $E_{\varepsilon}$ is over \SI{100}{\percent} for all models, indicating that this large increase in error is caused by the accumulation of floating point error from the forward and reverse ROPs, as previously discussed.
 The species rate error norm is similar in magnitude to that of the net ROP, however the temperature rate error again increases as it depends directly on the species rates.
-We note that the above discussion does not imply that computation of the net ROP will cause large error in chemical kinetic integration---either in this code or Cantera~\cite{Cantera}---as this loss of precision only occurs when the forward and reverse ROP are nearly equal, implying the reaction is in near-equilibrium.
-\todo{update for new valid}
+We note that the above discussion does not imply that computation of the net ROP will cause large error in chemical kinetic integration---either in this code or Cantera~\cite{Cantera}---as this loss of accuracy only occurs when the forward and reverse ROP are nearly equal, implying the reaction is in near-equilibrium.
 

### 2017-02-26 paper.tex
update wording

@@
 \end{equation}
-If this estimate is near \SI{100}{\percent}, it indicates that the maximum error in forward or reverse ROPs are of a similar magnitude to the net ROP, hence the error in net ROP would be expected to be large in this case.
+If this estimate is near \SI{100}{\percent}, it indicates that the maximum error in forward or reverse ROPs are of a similar magnitude to the net ROP, hence the error in net ROP is expected be large in this case.
 

### 2017-02-26 paper.tex
wording

@@
 \end{equation}
-and the error in reverse ROP $\varepsilon^{\prime\prime}_{ij}$ can be defined analogously.
+while the error in reverse ROP $\varepsilon^{\prime\prime}_{ij}$ can be defined analogously.
 Finally, for the reaction $i^{*}$ and state $j^{*}$ that result in the largest error in net ROP, i.e.
@@
 \end{equation}
-If this estimate is near \SI{100}{\percent}, it indicates that the maximum error in forward or reverse ROPs are of a similar magnitude to the net ROP, hence the error in net ROP is expected be large in this case.
+If this estimate is near \SI{100}{\percent}, it indicates that the error in forward or reverse ROPs at this state is of a similar magnitude to the net ROP, hence the error in net ROP is expected to be large.
 
@@
 We see here that $E_{\varepsilon}$ is over \SI{100}{\percent} for all models, indicating that this large increase in error is caused by the accumulation of floating point error from the forward and reverse ROPs, as previously discussed.
-The species rate error norm is similar in magnitude to that of the net ROP, however the temperature rate error again increases as it depends directly on the species rates.
+The species rate error norm is similar in magnitude to that of the net ROP, however the temperature rate error again increases as the error in species rates is amplified by multiplication with thermodynamic properties.
 We note that the above discussion does not imply that computation of the net ROP will cause large error in chemical kinetic integration---either in this code or Cantera~\cite{Cantera}---as this loss of accuracy only occurs when the forward and reverse ROP are nearly equal, implying the reaction is in near-equilibrium.

### 2017-02-27 scripts/error.py
update

@@ -16,7 +16,7 @@ for mech in mechs:
     for file in files:
         arrs = np.load(file)
         for name in arrs:
-            if 'value' in name or 'component' in name:
+            if 'value' in name or 'component' in name or 'store' in name:
                 continue
             errs = arrs[name]
             values = arrs[name + '_value']
@@ -25,7 +25,7 @@ for mech in mechs:
             precs = None
             if 'rop_net' in name:
                 #calculate the precision norms
-                precs = 100 * arrs['rop_component'] / arrs[name]
+                precs = 100 * arrs['rop_component'] / np.abs(values)
 
             if name in err_dicts[mech_name]:
                 err_dicts[mech_name][name] = np.maximum(err_dicts[mech_name][name], errs)
@@ -52,10 +52,11 @@ for mech in err_dicts:
         if 'wdot' in name:
             print('tdot', format(err_vals[0]))
             print('species', format(np.linalg.norm(err_vals[1:], ord=np.inf)))
-        elif 'rop_net ' in name:
+        elif 'rop_net' in name:
             #find prevision range
             maxv = np.linalg.norm(err_vals, ord=np.inf)
             maxind = np.where(err_dicts[mech][name] == maxv)[0][0]
             print(name, format(maxv))
-        elif 'prec' not in name:
+            print('rop_component', format(err_dicts[mech]['rop_component'][maxind]))
+        elif 'component' not in name:
             print(name, format(np.linalg.norm(err_vals, ord=np.inf)))

### 2017-02-27 paper.tex
comma
update the norm for clarity

@@
 \end{equation}
-while the error in reverse ROP $\varepsilon^{\prime\prime}_{ij}$ can be defined analogously.
+while the error in reverse ROP, $\varepsilon^{\prime\prime}_{ij}$, can be defined analogously.
 Finally, for the reaction $i^{*}$ and state $j^{*}$ that result in the largest error in net ROP, i.e.
@@
 \begin{equation}
-E_{\varepsilon} = 100\times\frac{\max(\varepsilon^{\prime}_{i^{*}j^{*}}\text{, }\varepsilon^{\prime\prime}_{i^{*}j^{*}})}{R_{i^{*}j^{*},\text{CT}}} \si{\percent}
+E_{\varepsilon} = \frac{\max(\varepsilon^{\prime}_{i^{*}j^{*}}\text{, }\varepsilon^{\prime\prime}_{i^{*}j^{*}})}{\num{e-10} + \num{e-6} * \left\lvert R_{i^{*}j^{*},\text{CT}} \right\rvert}
 \end{equation}
-If this estimate is near \SI{100}{\percent}, it indicates that the error in forward or reverse ROPs at this state is of a similar magnitude to the net ROP, hence the error in net ROP is expected to be large.
+This estimate allows for direct comparison of the error in forward or reverse ROPs to the net ROP, if they are of similar magnitude the error in net ROP will be large.
 
@@
 \toprule
-\multicolumn{1}{l}{Model} & \multicolumn{1}{c}{$E_{R^{\prime}}$} & \multicolumn{1}{c}{$E_{R^{\prime\prime}}$} &\multicolumn{1}{c}{$E_{R}$} & \multicolumn{1}{c}{$E_{\varepsilon}\si{\percent}$} & \multicolumn{1}{c}{$E_{\dot{\omega}}$} & \multicolumn{1}{c}{$E_{\frac{\partial T}{\partial t}}$} \\
+\multicolumn{1}{l}{Model} & \multicolumn{1}{c}{$E_{R^{\prime}}$} & \multicolumn{1}{c}{$E_{R^{\prime\prime}}$} &\multicolumn{1}{c}{$E_{R}$} & \multicolumn{1}{c}{$E_{\varepsilon}$} & \multicolumn{1}{c}{$E_{\dot{\omega}}$} & \multicolumn{1}{c}{$E_{\frac{\partial T}{\partial t}}$} \\
 \midrule
-\multicolumn{1}{l}{\ce{H2}\slash~\ce{CO}\cite{Burke:2011fh}} & 1.20e-8 & 6.33e-8 & 1.32e1 & 1.08e2 & 2.37e1 & 2.12e5 \\
-\multicolumn{1}{l}{GRI-Mech.~3.0\cite{smith_gri-mech_30}}  & 3.07e-8 & 5.90e-8 & 1.21e0 & 1.42e2 & 2.84e0 & 2.64e4 \\
-\multicolumn{1}{l}{USC-Mech II\cite{Wang:2007}}  & 1.07e-7 & 1.37e-7 & 2.39e0 & 4.12e02 & 1.44e2 & 3.22e03 \\
-\multicolumn{1}{l}{\ce{iC5H11OH}\cite{Sarathy:2013jr}} & 6.80e-8 &  1.48e-7 & 5.26e-1 & 1.59e2 & 2.26e00 & 6.10e+01 \\
+\multicolumn{1}{l}{\ce{H2}\slash~\ce{CO}\cite{Burke:2011fh}} & 1.20e-8 & 6.33e-8 & 1.32e1 & 1.32e1 & 2.37e1 & 2.12e5 \\
+\multicolumn{1}{l}{GRI-Mech.~3.0\cite{smith_gri-mech_30}}  & 3.07e-8 & 5.90e-8 & 1.21e0 & 1.29e90 & 2.84e0 & 2.64e4 \\
+\multicolumn{1}{l}{USC-Mech II\cite{Wang:2007}}  & 1.07e-7 & 1.37e-7 & 2.39e0 & 2.27e0 & 1.44e2 & 3.22e03 \\
+\multicolumn{1}{l}{\ce{iC5H11OH}\cite{Sarathy:2013jr}} & 6.80e-8 &  1.48e-7 & 5.26e-1 & 4.87e-1 & 2.26e00 & 6.10e+01 \\
 \bottomrule

### 2017-02-27 scripts/error.py

@@ -25,7 +25,7 @@ for mech in mechs:
             precs = None
             if 'rop_net' in name:
                 #calculate the precision norms
-                precs = 100 * arrs['rop_component'] / np.abs(values)
+                precs = arrs['rop_component'] / (atol + rtol * np.abs(values))
 
             if name in err_dicts[mech_name]:
                 err_dicts[mech_name][name] = np.maximum(err_dicts[mech_name][name], errs)

### 2017-02-27 paper.tex
wording

@@
 \end{equation}
-This estimate allows for direct comparison of the error in forward or reverse ROPs to the net ROP, if they are of similar magnitude the error in net ROP will be large.
+This estimate allows for direct comparison of the error in forward or reverse ROPs to the magnitude of the net ROP itself, if they are of similar magnitude the error in net ROP will be large.
 
@@
 \multicolumn{1}{l}{\ce{H2}\slash~\ce{CO}\cite{Burke:2011fh}} & 1.20e-8 & 6.33e-8 & 1.32e1 & 1.32e1 & 2.37e1 & 2.12e5 \\
-\multicolumn{1}{l}{GRI-Mech.~3.0\cite{smith_gri-mech_30}}  & 3.07e-8 & 5.90e-8 & 1.21e0 & 1.29e90 & 2.84e0 & 2.64e4 \\
+\multicolumn{1}{l}{GRI-Mech.~3.0\cite{smith_gri-mech_30}}  & 3.07e-8 & 5.90e-8 & 1.21e0 & 1.29e0 & 2.84e0 & 2.64e4 \\
 \multicolumn{1}{l}{USC-Mech II\cite{Wang:2007}}  & 1.07e-7 & 1.37e-7 & 2.39e0 & 2.27e0 & 1.44e2 & 3.22e03 \\

### 2017-02-27 paper.tex
wording updates

--- 2017-02-27 paper.tex
@@
 Finally, for the reaction $i^{*}$ and state $j^{*}$ that result in the largest error in net ROP, i.e.
-$E_{R}$, an estimate of the percent error attributable to floating point error accumulation from the forward and reverse ROPs can be obtained as:
+$E_{R}$, an estimate of the error attributable to floating point error accumulation from the forward and reverse ROPs can be obtained as:
 \begin{equation}
@@
 \end{equation}
-This estimate allows for direct comparison of the error in forward or reverse ROPs to the net ROP, if they are of similar magnitude the error in net ROP will be large.
+This estimate allows for direct comparison of the error in forward or reverse ROPs to the value of the net ROP itself, if they are of similar magnitude the error in net ROP will be large.
 
@@
 \multicolumn{1}{l}{\ce{H2}\slash~\ce{CO}\cite{Burke:2011fh}} & 1.20e-8 & 6.33e-8 & 1.32e1 & 1.32e1 & 2.37e1 & 2.12e5 \\
-\multicolumn{1}{l}{GRI-Mech.~3.0\cite{smith_gri-mech_30}}  & 3.07e-8 & 5.90e-8 & 1.21e0 & 1.29e90 & 2.84e0 & 2.64e4 \\
+\multicolumn{1}{l}{GRI-Mech.~3.0\cite{smith_gri-mech_30}}  & 3.11e-8 & 5.90e-8 & 1.21e0 & 1.29e0 & 2.84e0 & 2.64e4 \\
 \multicolumn{1}{l}{USC-Mech II\cite{Wang:2007}}  & 1.07e-7 & 1.37e-7 & 2.39e0 & 2.27e0 & 1.44e2 & 3.22e03 \\
@@
 In Table~\ref{T:error}, we see the results of this code as compared to Cantera~\cite{Cantera} on a library of partially stirred reaction conditions (PaSR) described in our previous works~\cite{CurtisGPU:2017,Niemeyer:2016aa}.
-Very close agreement is observed for the forward and reverse ROP for all models, however the net ROP error norm is \numrange{6}{9} orders of magnitude larger.
-We see here that $E_{\varepsilon}$ is over \SI{100}{\percent} for all models, indicating that this large increase in error is caused by the accumulation of floating point error from the forward and reverse ROPs, as previously discussed.
+Very close agreement is observed for the forward and reverse ROP for all models, however the net ROP error norm is \numrange{7}{9} orders of magnitude larger.
+We see here that $E_{\varepsilon}$ is very similar in magnitude to $E_R$ in all cases, indicating that this large increase in error is caused by the accumulation of floating point error from the forward and reverse ROPs, as previously discussed.
 The species rate error norm is similar in magnitude to that of the net ROP, however the temperature rate error again increases as the error in species rates is amplified by multiplication with thermodynamic properties.

### 2017-02-27 scripts/error.py

@@ -42,7 +42,8 @@ def format(val):
     return '{:1.2e}'.format(val)
 
 keyarr = ['fwd', 'rev', 'net', 'comp', 'wdot']
-for mech in err_dicts:
+mech_arr = ['H2', 'CH4', 'C2H4', 'IC5H11OH']
+for mech in sorted(err_dicts, key=lambda x:mech_arr.index(next(y for y in mech_arr if y in x))):
     print(mech)
     for name in sorted(err_dicts[mech], key=lambda x:keyarr.index(next(y for y in keyarr if y in x))):
         if 'l2' in name:
